---
title: "Human Macrophages: M-CSF v GM-CSF: LPS, LPS+Adenosine, LPS+PGE2"
author: "Kajal Hamidzadeh and atb"
date: "`r Sys.Date()`"
output:
  html_document:
    code_download: true
    code_folding: show
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: tango
    keep_md: false
    mode: selfcontained
    number_sections: true
    self_contained: true
    theme: readable
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
      rmdformats::readthedown:
        code_download: true
        code_folding: show
        df_print: paged
        fig_caption: true
        fig_height: 7
        fig_width: 7
        highlight: tango
        width: 300
        keep_md: false
        mode: selfcontained
        toc_float: true
        BiocStyle::html_document:
          code_download: true
          code_folding: show
          fig_caption: true
          fig_height: 7
          fig_width: 7
          highlight: tango
          keep_md: false
          mode: selfcontained
          toc_float: true
---

<style type="text/css">
body, td {
  font-size: 16px;
}
code.r{
  font-size: 16px;
}
pre {
  font-size: 16px
}
</style>

```{r options, include=FALSE}
library("hpgltools")
tt <- devtools::load_all("/data/hpgltools")
knitr::opts_knit$set(width=120,
                     progress=TRUE,
                     verbose=TRUE,
                     echo=TRUE)
knitr::opts_chunk$set(error=TRUE,
                      dpi=96)
old_options <- options(digits=4,
                       stringsAsFactors=FALSE,
                       knitr.duplicate.label="allow")
ggplot2::theme_set(ggplot2::theme_bw(base_size=10))
rundate <- format(Sys.Date(), format="%Y%m%d")
previous_file <- ""
ver <- "20200330"

## tmp <- sm(loadme(filename=paste0(gsub(pattern="\\.Rmd", replace="", x=previous_file), "-v", ver, ".rda.xz")))
## rmd_file <- "03_expression_infection_20180822.Rmd"
```

# A version of Kajal's R markdown document with some more minor changes

## Libraries

```{r libraries, warning=FALSE, message=FALSE, include=FALSE}
library(devtools)
library(tximport)
library(biomaRt)
library(hpgltools)
library(DESeq2)
library(gplots)
library(ggplot2)
library(cbcbSEQ)
library(RColorBrewer)
library(Vennerable)
library(edgeR)
library(calibrate)
library(scales)
```

In this version of Kajal's document, I will repeat each step with some changes
to see if I can make things a little easier for future modification.

I will therefore leave Kajal's work unchanged and follow each block with an alternative.

## Set working directory, import metadata and abundance files

```{r read_data}
design <- read.table("sample_sheets/MetaData only 4 hour.txt", header=TRUE, sep="\t")
design[["Patient"]] <- as.factor(design[["Patient"]])
design[["Stimulation"]] <- as.factor(design[["Stimulation"]])
design[["Batch"]] <- as.factor(design[["Batch"]])
design[["Growth"]] <- as.factor(design[["Growth"]])
files <- file.path("kallisto abundance files/", design$HPGL.Identifier, "abundance.tsv")
names(files) <- paste0("HPGL09", c(12:31, 42:60))
rownames(design) <- design[[1]]

stim_design_idx <- design[["Stimulation"]] != "NS"
stim_design <- design[stim_design_idx, ]
```

## Convert transcript ID to gene ID

```{r ensembl}
ensembl <- useMart("ENSEMBL_MART_ENSEMBL", dataset="hsapiens_gene_ensembl",
                   host="useast.ensembl.org")
tx2gene <- getBM(attributes=c("ensembl_transcript_id", "ensembl_gene_id", "chromosome_name"),
                 mart=ensembl)
good_idx <- grepl(x=tx2gene[["chromosome_name"]], pattern="^[[:alnum:]]{1,2}$")
good_ones <- tx2gene[good_idx, -3]
```

## Create counts table

```{r}
txi.kallisto.tsv <- tximport(files, type="kallisto", tx2gene=good_ones,
                             countsFromAbundance="lengthScaledTPM")
nrow(txi.kallisto.tsv$counts)

write.table(txi.kallisto.tsv$counts, "TPM_MvGM_CDS_20180711.txt",
            col.names=TRUE, row.names=FALSE, quote=FALSE)
write.csv(txi.kallisto.tsv$counts, "csv/TPM_MvGM_20180711.csv",
          row.names=TRUE, quote=FALSE)

stim_counts <- txi.kallisto.tsv$counts
stim_counts_idx <- colnames(stim_counts) %in% rownames(stim_design)
stim_counts <- stim_counts[, stim_counts_idx]
```

This is a potentially important difference in how I and Kajal treated the data.
This is because I created the expressionset without using 'lengthScaledTPM',
but let tximport use the raw counts so that the various normalizations are not
affected.  E.g. using these scaled numbers results in less variance in the data
and may lead to some confusion for the downstream tools
(edger/limma/deseq/etc).  On the other hand, it may lead to cleaner plots, I am
not sure.

## Create DESeqDataSet

Not sure what this is actually doing

```{r deseq_dds}
df <- data.frame(condition=paste(design$HPGL.Identifier, design$Growth,
                                 design$Stimulation, design$Patient, sep="_"))
rownames(df) <- colnames(txi.kallisto.tsv$counts)
dds <- DESeqDataSetFromTximport(txi.kallisto.tsv, df, ~condition)
nrow(dds)

dds_stim <- dds[, rownames(stim_design)]
```

## My version of the above

```{r atb_read_data}
my_design <- design
rownames(my_design) <- design[[1]]
my_design[["condition"]] <- my_design[["Stimulation"]]
my_design[["file"]] <- glue::glue("preprocessing/{rownames(my_design)}/abundance.tsv")
colnames(my_design) <- tolower(colnames(my_design))
gene_info <- load_biomart_annotations(host="useast.ensembl.org")$annotation
rownames(gene_info) <- make.names(gene_info[["ensembl_gene_id"]], unique=TRUE)
tx_gene_map <- gene_info[, c("ensembl_transcript_id", "ensembl_gene_id")]
hs_expt <- create_expt(metadata=my_design, gene_info=gene_info, tx_gene_map=tx_gene_map)
hs_expt <- set_expt_batches(hs_expt, fact="growth")

stim_expt <- subset_expt(hs_expt, subset="stimulation!='NS'")
```

## Bar Plot of Counts

```{r barplot}
par(mar=c(10, 4.5, 3.5, 1))
par(oma=c(0, 0, 0, 0))
barplot(colSums(txi.kallisto.tsv$counts), las=3, main="Raw Counts By Sample")
barplot(colSums(stim_counts), las=3, main="Raw Counts By Sample")
```

## My Barplot of counts

```{r my_barplot}
libsize <- plot_libsize(hs_expt)
libsize$plot
libsize$summary

plot_libsize(stim_expt)$plot
```

## Box Plot of Counts

```{r count_barplot}
dds <- estimateSizeFactors(dds)
ncts <- counts(dds, normalized=TRUE)
y <- log(ncts + 1)

par(mar=c(5, 5, 2, 1))
par(oma=c(0, 0, 0, 0))
boxplot(y, names=colnames(txi.kallisto.tsv$counts),
        las=3, main="Per Sample Log of Size-factor Counts")

dds_stim <- estimateSizeFactors(dds_stim)
stim_ncts <- counts(dds_stim, normalized=TRUE)
stim_y <- log(stim_ncts + 1)

par(mar=c(5, 5, 2, 1))
par(oma=c(0, 0, 0, 0))
boxplot(stim_y, names=colnames(stim_counts), las=3,
        main="Per Sample Log of Size-factor Counts")
```

```{r my_boxplot}
boxplot <- plot_boxplot(hs_expt)
boxplot
```

## Heatmap of Pearson Correlation

```{r corheat}
datCor <- cor(txi.kallisto.tsv$counts)
heatmap.2(datCor, Rowv=NA, Colv=NA, margins=c(10, 10),
          labRow=df$condition, labCol=df$condition,
          dendrogram="none", scale="none", trace="none",
          srtCol=45, main="Pearson Correlation")

stim_datCor <- cor(stim_counts)
heatmap.2(stim_datCor, Rowv=NA, Colv=NA, margins=c(10, 10),
          labRow=df$condition, labCol=df$condition,
          dendrogram="none", scale="none", trace="none",
          srtCol=45, main="Pearson Correlation")
```

### Try out ggstatsplot for this

```{r ggtesting}
library(ggstatsplot)
ggtest <- ggcorrmat(as.data.frame(stim_counts), type="robust",
                    sig.level=0.001, p.adjust.method="holm")
limits <- c(min(ggtest[["data"]][["coef"]]), 1)
ggtest <- ggtest + scale_fill_gradient(
             low="black", high="red", limit=limits) +
  theme(axis.text.y=element_text(size=6),
        legend.position="top")

library(ggdendro)
dendro_mtrx <- dcast(data=ggtest[["data"]], formula=Var1 ~ Var2, value.var="coef")
rownames(dendro_mtrx) <- dendro_mtrx[["Var1"]]
dendro_mtrx[["Var1"]] <- NULL
dendro_mtrx <- as.matrix(dendro_mtrx)
dendro <- as.dendrogram(hclust(d=dist(x=dendro_mtrx)))
dendro_plot <- ggdendrogram(data=dendro, rotate=TRUE)

library(grid)
grid.newpage()
print(ggtest, vp = viewport(x = 0.4, y = 0.5, width = 0.8, height = 1.0))
print(dendro_plot, vp = viewport(x = 0.90, y = 0.445, width = 0.2, height = 1.0))
```


```{r my_corheat}
plot_corheat(hs_expt)$plot
plot_corheat(stim_expt)$plot
```

## Filter and Normalize Counts

```{r}
filterCounts <- function (counts, lib.size=NULL, thresh=1, minSamples=2) {
  cpms <- 2 ^ log2CPM(counts, lib.size = lib.size)$y
  keep <- rowSums(cpms > thresh) >= minSamples
  counts <- counts[keep, ]
  counts
}

x_table <- table(design$Stimulation)
dim(txi.kallisto.tsv$counts)

counts <- filterCounts(txi.kallisto.tsv$counts, thresh=1, minSamples=min(x_table))
dim(counts)
countsSubQ <- qNorm(counts)
x <- log2CPM(countsSubQ)
s <- makeSVD(x$y)

stim_table <- table(stim_design$Stimulation)
stim_counts <- filterCounts(stim_counts, thresh=1, minSamples=min(stim_table))
dim(stim_counts)
stim_countsSubQ <- qNorm(stim_counts)
stim_x <- log2CPM(stim_countsSubQ)
stim_s <- makeSVD(stim_x$y)
```

## Median Pairwise Correlation

```{r smc}
corM <- matrixStats::rowMedians(cor(txi.kallisto.tsv$counts))
qs <- quantile(corM, p=c(1, 3) / 4)
iqr <- diff(qs)
outLimit <- qs[1] - 1.5 * iqr
ylim <- c(pmin(min(corM), outLimit), max(corM))
cond <- paste(design$Growth, design$Stimulation, sep="_")
col <- ifelse(cond == "M_NS", "gray60",
       ifelse(cond == "M_LPS", "deeppink3",
       ifelse(cond == "M_LA", "darkseagreen2",
       ifelse(cond == "M_LP", "lavender",
       ifelse(cond == "GM_NS", "lightpink1",
       ifelse(cond == "GM_LPS", "moccasin",
       ifelse(cond == "GM_LA", "black",
       ifelse(cond == "GM_LP", "springgreen4", "blue2"))))))))
par(mar=c(5, 4.5, 2, 1))
plot(corM, xaxt="n", ylim=ylim, ylab="Median Pairwise Correlation",
     xlab="", main="", col=col, pch=16, cex=1.5)
axis(side=1, at=seq(along=corM), labels=colnames(txi.kallisto.tsv$counts), las=2)
abline(h=outLimit, lty=2)
abline(v=1:ncol(txi.kallisto.tsv$counts), lty=3, col="black")

stim_corM <- matrixStats::rowMedians(cor(stim_counts))
stim_qs <- quantile(stim_corM, p=c(1, 3) / 4)
stim_iqr <- diff(stim_qs)
stim_outLimit <- stim_qs[1] - (1.5 * iqr)
stim_ylim <- c(pmin(min(stim_corM), stim_outLimit), max(stim_corM))
stim_cond <- paste(stim_design$Growth, stim_design$Stimulation, sep="_")
col <- ifelse(cond=="M_NS", "gray60",
       ifelse(cond=="M_LPS", "deeppink3",
       ifelse(cond=="M_LA", "darkseagreen2",
       ifelse(cond=="M_LP", "lavender",
       ifelse(cond=="GM_NS", "lightpink1",
       ifelse(cond=="GM_LPS", "moccasin",
       ifelse(cond=="GM_LA", "black",
       ifelse(cond=="GM_LP", "springgreen4", "blue2"))))))))
par(mar=c(5, 4.5, 2, 1))
plot(stim_corM, xaxt="n", ylim=ylim, ylab="Median Pairwise Correlation",
     xlab="", main="", col=col, pch=16, cex=1.5)
axis(side=1, at=seq(along=stim_corM), labels=colnames(stim_counts), las=2)
abline(h=stim_outLimit, lty=2)
abline(v=1:ncol(stim_counts), lty=3, col="black")
```

```{r my_smc}
plot_sm(hs_expt)$plot
plot_sm(stim_expt)$plot
```

## PCA

```{r}
cbcbSEQ::pcRes(s$v, s$d, cond, design$Batch)[1:5, ]
cbcbSEQ::pcRes(s$v, s$d, cond, design$Patient)[1:5, ]
cbcbSEQ::pcRes(stim_s$v, stim_s$d, stim_cond, stim_design$Batch)[1:5, ]
cbcbSEQ::pcRes(stim_s$v, stim_s$d, stim_cond, stim_design$Patient)[1:5, ]
```

## Euclidian Distance Heat Map

```{r disheat}
dists <- dist(t(counts))
mat <- as.matrix(dists)
rownames(mat) <- colnames(mat) <- cond
hmcol <- colorRampPalette(brewer.pal(9, "GnBu"))(100)
vec.patient <- rainbow(nlevels(design$Patient), start=0, end=0.8)
patient.color <- rep(0, length(design$Patient))
for (i in 1:length(design$Patient)) {
  patient.color[i] <- vec.patient[design$Patient[i] == levels(design$Patient)]
}
vec.condition <- c("green", "lightblue", "pink", "purple")
condition.color <- rep(0, length(design$Stimulation))
for (i in 1:length(design$Stimulation)) {
  condition.color[i] <- vec.condition[design$Stimulation[i] == levels(design$Stimulation)]
}

heatmap <- heatmap.2(mat, trace="none", col=rev(hmcol), margin=c(11, 11),
                     ColSideColors=condition.color,
                     RowSideColors=patient.color, key=FALSE, srtCol=45)

stim_dists <- dist(t(stim_counts))
stim_mat <- as.matrix(stim_dists)
rownames(stim_mat) <- colnames(stim_mat) <- stim_cond
stim_hmcol <- colorRampPalette(brewer.pal(9, "GnBu"))(100)
stim_vec.patient <- rainbow(nlevels(stim_design$Patient), start=0, end=0.8)
stim_patient.color <- rep(0, length(stim_design$Patient))
for (i in 1:length(stim_design$Patient)) {
  idx <- stim_design$Patient[i] == levels(stim_design$Patient)
  stim_patient.color[i] <- stim_vec.patient[idx]
}
stim_vec.condition <- c("green", "lightblue", "pink", "purple")
stim_condition.color <- rep(0, length(stim_design$Stimulation))
for (i in 1:length(stim_design$Stimulation)) {
  idx <- stim_design$Stimulation[i] == levels(stim_design$Stimulation)
  stim_condition.color[i] <- stim_vec.condition[idx]
}
heatmap <- heatmap.2(stim_mat, trace="none", col=rev(hmcol), margin=c(11, 11),
                     ColSideColors=stim_condition.color,
                     RowSideColors=stim_patient.color, key=FALSE, srtCol=45)
```

```{r my_disheat}
my_disheat <- plot_disheat(hs_expt)
stim_disheat <- plot_disheat(stim_expt)
```

## Plot PC1 v PC2

```{r original_pca}
condnum <- as.numeric(as.factor(design$Stimulation))
condnum <- ifelse(condnum == 4, "green",
           ifelse(condnum == 3, "lightblue",
           ifelse(condnum == 2, "pink",
           ifelse(condnum == 1, "purple", "black"))))
patnum <- as.numeric(as.factor(design$Patient))
patnum <- ifelse(patnum == 4, 21,
          ifelse(patnum == 3, 22,
          ifelse(patnum == 2, 23,
          ifelse(patnum == 1, 24,
          ifelse(patnum == 5, 25, 1)))))
cbcbSEQ::plotPC(s$v, s$d, col="black", pch=patnum, bg=condnum)
legend(x=0.2, y=0.2, legend=unique(design$Stimulation), pch=22, col=0,
       pt.bg=c("green", "lightblue", "pink", "purple"), pt.cex=1.5, cex=0.5, bty="n")
legend(x=0.2, y=-0.1, legend=unique(design$Patient), pch=unique(patnum), col=0,
       pt.bg="gray90", pt.cex=1.5, cex=0.5, bty="n")

stim_condnum <- as.numeric(as.factor(stim_design$Stimulation))
stim_condnum <- ifelse(stim_condnum == 4, "green",
                ifelse(stim_condnum == 3, "lightblue",
                ifelse(stim_condnum == 2, "pink",
                ifelse(stim_condnum == 1, "purple", "black"))))
stim_patnum <- as.numeric(as.factor(stim_design$Patient))
stim_patnum <- ifelse(stim_patnum == 4, 21,
               ifelse(stim_patnum == 3, 22,
               ifelse(stim_patnum == 2, 23,
               ifelse(stim_patnum == 1, 24,
               ifelse(stim_patnum == 5, 25, 1)))))
cbcbSEQ::plotPC(stim_s$v, stim_s$d, col="black", pch=stim_patnum, bg=stim_condnum)
legend(x=0.2, y=0.2, legend=unique(stim_design$Stimulation), pch=22, col=0,
       pt.bg=c("green","lightblue","pink","purple"), pt.cex=1.5, cex=0.5, bty="n")
legend(x=0.2, y=-0.1, legend=unique(stim_design$Patient), pch=unique(stim_patnum), col=0,
       pt.bg="gray90", pt.cex=1.5, cex=0.5, bty="n")
```

```{r my_pca}
hs_norm <- normalize_expt(hs_expt, norm="quant", filter="cbcb", transform="log2")
my_pca <- plot_pca(hs_norm, plot_labels=FALSE, cis=NULL)
my_pca$plot

hsstim_norm <- normalize_expt(stim_expt, norm="quant", filter="cbcb", transform="log2")
stim_pca <- plot_pca(hsstim_norm, plot_labels=FALSE, cis=NULL)
stim_pca$plot
```

## Correct for Patient in Limma model

```{r}
mod <- model.matrix(~design$Patient)
v <- voom(countsSubQ, mod)
fit <- lmFit(v)
newData <- residuals(fit, v)
s <- makeSVD(newData)
pcRes(s$v, s$d, cond, design$Batch)[1:5, ]
pcRes(s$v, s$d, cond, design$Patient)[1:5, ]

stim_table <- table(stim_design$Stimulation)
stim_counts <- filterCounts(stim_counts, thresh=2, minSamples=min(stim_table))
dim(stim_counts)
stim_countsSubQ <- qNorm(stim_counts)
stim_mod <- model.matrix(~stim_design$Patient)
stim_v <- voom(stim_countsSubQ, stim_mod)
stim_fit <- lmFit(stim_v)
stim_newData <- residuals(stim_fit, stim_v)
stim_s <- makeSVD(stim_newData)
pcRes(stim_s$v, stim_s$d, stim_cond, stim_design$Batch)[1:5, ]
pcRes(stim_s$v, stim_s$d, stim_cond, stim_design$Patient)[1:5, ]
```

## Plot PC1 and PC2 with patient correction

```{r}
gronum <- as.numeric(as.factor(design$Growth))
gronum <- ifelse(gronum == 2, 19,
          ifelse(gronum == 1, 15, 1))
patnum <- as.numeric(as.factor(design$Patient))
patnum <- ifelse(patnum == 1, "pink",
          ifelse(patnum == 2, "green",
          ifelse(patnum == 3, "blue",
          ifelse(patnum == 4, "yellow",
          ifelse(patnum == 5, "black", "orange")))))
samplenum <- as.numeric(as.factor(design$Stimulation))
samplenum <- ifelse(samplenum == 4, "green",
             ifelse(samplenum == 3, "blue",
             ifelse(samplenum == 1, "black",
             ifelse(samplenum == 2, "yellow", "grey"))))
plotPC(s$v, s$d, pch=gronum, col=samplenum, cex=2)
legend(x=0.05, y=0.3, legend=c("NS", "LPS", "LP", "LA"), pch=22, col=0,
       pt.bg=c("green", "blue", "black", "yellow"), pt.cex=1.5, cex=1, bty="n")
legend(x=0.05, y=0.1, legend=unique(design$Growth), pch=unique(gronum), col="black",
       pt.bg="black", pt.cex=1.0, cex=1, bty="n")

stim_gronum <- as.numeric(as.factor(stim_design$Growth))
stim_gronum <- ifelse(stim_gronum == 2, 19,
               ifelse(stim_gronum == 1, 15, 1))
stim_patnum <- as.numeric(as.factor(stim_design$Patient))
stim_patnum <- ifelse(stim_patnum == 1, "pink",
               ifelse(stim_patnum == 2, "green",
               ifelse(stim_patnum == 3, "blue",
               ifelse(stim_patnum == 4, "yellow",
               ifelse(stim_patnum == 5, "black", "orange")))))
stim_samplenum <- as.numeric(as.factor(stim_design$Stimulation))
stim_samplenum <- ifelse(stim_samplenum == 3, "green",
                  ifelse(stim_samplenum == 2, "blue",
                  ifelse(stim_samplenum == 1, "black", "grey")))
plotPC(stim_s$v, stim_s$d, pch=stim_gronum, col=stim_samplenum, cex=2)
legend(x=0.05, y=0.3, legend=c("LPS", "LP", "LA"), pch=22, col=0,
       pt.bg=c("green", "blue", "black", "yellow"), pt.cex=1.5, cex=1, bty="n")
legend(x=0.05, y=0.1, legend=unique(stim_design$Growth), pch=unique(stim_gronum),
       col="black", pt.bg="black", pt.cex=1.0, cex=1, bty="n")
```

## My Figure 3A

```{r my_pca2}
##stim_counts <- filterCounts(stim_counts, thresh=1, minSamples=min(x))
##dim(stim_counts)
##stim_countsSubQ <- qNorm(stim_counts)
##stim_x <- log2CPM(stim_countsSubQ)
##stim_s <- makeSVD(stim_x$y)
hss <- set_expt_batches(stim_expt, fact="patient")
hss <- sm(normalize_expt(hss, filter=TRUE))
hss <- sm(normalize_expt(hss, norm="quant"))
hss <- sm(normalize_expt(hss, convert="cpm"))
hss <- sm(normalize_expt(hss, transform="log2"))
hss <- sm(normalize_expt(hss, batch="limma"))
hss <- sm(set_expt_batches(hss, fact="growth"))
plot_pca(hss, plot_labels=FALSE)$plot

hss_varpart <- simple_varpart(expt=hss)
hss_varplot <- plot_variance_coefficients(hss)
pp(file="stimulation_coef.png", image=hss_varplot$plot)

hss_varplot <- plot_variance_coefficients(hss, x_axis="batch")
hss_varplot$plot
pp(file="mgm_coef.png", image=hss_varplot$plot)

tmp <- pData(hss)
tmp[["both"]] <- paste0(tmp[["condition"]], "_", tmp[["batch"]])
pData(hss[["expressionset"]]) <- tmp
hss_both_varplot <- plot_variance_coefficients(hss, x_axis="both")
pp(file="both_coef.png", image=hss_both_varplot$plot)
```

## Heatmap with patient correction??

```{r}
dists <- dist(t(newData))
mat <- as.matrix(dists)
rownames(mat) <- colnames(mat) <- cond
heatmap <- heatmap.2(mat, trace="none", col=rev(hmcol), margin=c(11, 11),
                     ColSideColors=condition.color,
                     RowSideColors=patient.color, key=FALSE, srtCol=45)
```

## DE analysis

```{r}
countsSubQ <- qNorm(counts)
patient <- design$Patient
mod <- model.matrix(~ 0 + cond + patient, data=design)
v <- voom(countsSubQ, mod, plot=TRUE)
fit <- lmFit(v)
```

## M_LPS v M_NS

eBayes finds an F-statistic from the set of t-statistics for that gene

```{r}
M_LPS.M_NS.contr.mat <- makeContrasts(M_LPSvM_NS=(condM_LPS - condM_NS), levels=v$design)
M_LPS.M_NS.fit <- contrasts.fit(fit, M_LPS.M_NS.contr.mat)
M_LPS.M_NS.eb <- eBayes(M_LPS.M_NS.fit)
M_LPS.M_NS.topTab <- topTable(M_LPS.M_NS.eb, coef="M_LPSvM_NS", number=nrow(v$E))
M_LPS.M_NS.topTab <- cbind(rownames(M_LPS.M_NS.topTab), M_LPS.M_NS.topTab)
colnames(M_LPS.M_NS.topTab) <- c("ID", "logFC", "AveExpr", "t", "P.Value", "adj.P.Val", "B")
rownames(M_LPS.M_NS.topTab) <- c(1:nrow(M_LPS.M_NS.topTab))

lps_ns_table <- "csv/topTab_M_LPSvM_NS_CDS_limmabatchcorrection_20171120.csv"
write.csv(M_LPS.M_NS.topTab, file=lps_ns_table,
          row.names=FALSE, quote=FALSE)
```

Limit list to genes with an adjusted p value < 0.05

```{r}
M_LPS.M_NS.sigGenes <- M_LPS.M_NS.topTab[M_LPS.M_NS.topTab$adj.P.Val < 0.05, ]
length(M_LPS.M_NS.sigGenes$ID)
```

Filter out rows with less than 2-fold change (log2 fold change of > 1)

```{r}
M_LPS.M_NS.sigGenesFold1 <- subset(M_LPS.M_NS.sigGenes, abs(logFC) > 1)
length(M_LPS.M_NS.sigGenesFold1$ID)
```

Filter out rows with less than 4-fold change (log2 fold change of > 2)

```{r}
M_LPS.M_NS.sigGenesFold2 <- subset(M_LPS.M_NS.sigGenes, abs(logFC) > 2)
length(M_LPS.M_NS.sigGenesFold2$ID)
```

Make an MA plot

```{r}
sel <- M_LPS.M_NS.topTab$adj.P.Val < 0.05
top <- M_LPS.M_NS.topTab
sub <- paste("No. of sig. genes: ", sum(sel), "/", length(sel))
cpm <- v$E

plot(rowMeans(cpm[top$ID,]), top$logFC, pch=16, cex=0.5, col="darkgrey",
     main="M_LPSvM_NS adjusted", ylab="log FC",
     xlab="Average Expression", sub=sub)
points(rowMeans(cpm[top$ID, ])[sel], top$logFC[sel], col="red", cex=0.5)
abline(h=c(-1, 0, 1), col="red")
```

Annotate sigGenes list using Biomart

```{r}
M_LPS.M_NS.sigGenes <- M_LPS.M_NS.sigGenes[order(-M_LPS.M_NS.sigGenes$logFC), ]
sigGenes <- M_LPS.M_NS.sigGenes
ids <- sigGenes$ID

desc <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol", "description", "gene_biotype"),
              filters="ensembl_gene_id", values=ids, mart=ensembl)
colnames(desc) <- c("ID", "Symbol", "Description", "Type")
desc$Description <- gsub(",", "", desc$Description)

DEG_LPS <- merge(sigGenes, desc, by="ID", all=TRUE)
DEG_LPS <- subset(DEG_LPS,
                  select=c(ID, Symbol, Description, logFC, adj.P.Val, AveExpr, Type),
                  FS="/t")
DEG_LPS <- DEG_LPS[order(-DEG_LPS$logFC), ]
## Filter out Genes -1<FC<1
DEG_LPS <- subset(DEG_LPS, abs(DEG_LPS$logFC)>1)

## Save DE genes
write.table(DEG_LPS, "csv/DEG_M_LPSvM_NS_CDS_limmabatchcorrection_20171120.txt",
            col.names=TRUE, row.names=FALSE, quote=FALSE)
write.csv(DEG_LPS, "csv/DEG_M_LPSvM_NS_CDS_limmabatchcorrection_20171120.csv",
          row.names=FALSE, quote=FALSE)
```

## M_LPS v M_LA

```{r}
#eBayes finds an F-statistic from the set of t-statistics for that gene
M_LPS.M_LA.contr.mat <- makeContrasts(M_LPSvM_LA=((condM_LA - condM_NS) - (condM_LPS - condM_NS)),
                                      levels=v$design)
M_LPS.M_LA.fit <- contrasts.fit(fit, M_LPS.M_LA.contr.mat)
M_LPS.M_LA.eb <- eBayes(M_LPS.M_LA.fit)
M_LPS.M_LA.topTab <- topTable(M_LPS.M_LA.eb, coef="M_LPSvM_LA", number=nrow(v$E))
M_LPS.M_LA.topTab <- cbind(rownames(M_LPS.M_LA.topTab), M_LPS.M_LA.topTab)
colnames(M_LPS.M_LA.topTab) <- c("ID", "logFC", "AveExpr", "t", "P.Value", "adj.P.Val", "B")
rownames(M_LPS.M_LA.topTab) <- c(1:nrow(M_LPS.M_LA.topTab))

lps_la_table <- "csv/topTab_M_LPSvM_LA_CDS_limmabatchcorrection_20171120rev.csv"
write.csv(M_LPS.M_LA.topTab, file=lps_la_table, row.names=FALSE, quote=FALSE)

##Limit list to genes with an adjusted p value < 0.05
M_LPS.M_LA.sigGenes <- M_LPS.M_LA.topTab[M_LPS.M_LA.topTab$adj.P.Val < 0.05, ]
length(M_LPS.M_LA.sigGenes$ID)
##[1] 617

##Filter out rows with less than 2-fold change (log2 fold change of > 1)
M_LPS.M_LA.sigGenesFold1 <- subset(M_LPS.M_LA.sigGenes, abs(logFC) > 1)
length(M_LPS.M_LA.sigGenesFold1$ID)
##[1] 240

##Filter out rows with less than 4-fold change (log2 fold change of > 2)
M_LPS.M_LA.sigGenesFold2 <- subset(M_LPS.M_LA.sigGenes, abs(logFC) > 2)
length(M_LPS.M_LA.sigGenesFold2$ID)
##[1] 69

##Make an MA plot
sel <- M_LPS.M_LA.topTab$adj.P.Val < 0.05
top <- M_LPS.M_LA.topTab
sub <- paste("No. of sig. genes: ", sum(sel), "/", length(sel))
cpm <- v$E

plot(rowMeans(cpm[top$ID,]), top$logFC, pch=16, cex=0.5, col="darkgrey",
     main="M_LPSvM_LA adjusted",
     ylab="log FC", xlab="Average Expression",
     sub=sub)
points(rowMeans(cpm[top$ID, ])[sel], top$logFC[sel], col="red", cex=0.5)
abline(h=c(-1, 0, 1), col="red")

dev.copy(pdf, "images/MAplot_M_LPSvM_LA_CDS_limmabatchcorrection_20171120rev.pdf",
         width=8, height=8)
dev.off()
dev.copy(png, "images/MAplot_M_LPSvM_LA_CDS_limmabatchcorrection_20171120rev.png",
         width=700, height=700)
dev.off()

M_LPS.M_LA.sigGenes <- M_LPS.M_LA.sigGenes[order(-M_LPS.M_LA.sigGenes$logFC), ]

##Annotate sigGenes list using Biomart
sigGenes <- M_LPS.M_LA.sigGenes
ids <- sigGenes$ID

desc <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol", "description", "gene_biotype"),
              filters="ensembl_gene_id",
              values=ids, mart=ensembl)
colnames(desc) <- c("ID", "Symbol", "Description", "Type")

##Remove commas from description
desc$Description <- gsub(",", "", desc$Description)

DEG <- merge(sigGenes, desc, by="ID", all=TRUE)
DEG <- subset(DEG, select=c(ID, Symbol, Description, logFC, adj.P.Val, AveExpr, Type), FS="/t")
DEG <- DEG[order(-DEG$logFC), ]

##Filter out Genes -1<FC<1
DEG <- subset(DEG, abs(DEG$logFC) > 1)

##Save DE genes
write.table(DEG, "csv/DEG_M_LPSvM_LA_CDS_limmabatchcorrection_20171120rev.txt",
            col.names=TRUE, row.names=FALSE, quote=FALSE)
write.csv(DEG, "csv/DEG_M_LPSvM_LA_CDS_limmabatchcorrection_20171120rev.csv",
          row.names=FALSE, quote=FALSE)
```

## M_LPS v M_LP

```{r}
#eBayes finds an F-statistic from the set of t-statistics for that gene
M_LPS.M_LP.contr.mat <- makeContrasts(
    M_LPSvM_LP=((condM_LP - condM_NS) - (condM_LPS - condM_NS)),
    levels=v$design)
M_LPS.M_LP.fit <- contrasts.fit(fit, M_LPS.M_LP.contr.mat)
M_LPS.M_LP.eb <- eBayes(M_LPS.M_LP.fit)
M_LPS.M_LP.topTab <- topTable(M_LPS.M_LP.eb, coef="M_LPSvM_LP", number=nrow(v$E))
M_LPS.M_LP.topTab <- cbind(rownames(M_LPS.M_LP.topTab), M_LPS.M_LP.topTab)
colnames(M_LPS.M_LP.topTab) <- c("ID", "logFC", "AveExpr", "t", "P.Value", "adj.P.Val", "B")
rownames(M_LPS.M_LP.topTab) <- c(1:nrow(M_LPS.M_LP.topTab))

lps_lp_table <- "csv/topTab_M_LPSvM_LP_CDS_limmabatchcorrection_20171120rev.csv"
write.csv(M_LPS.M_LP.topTab, file=lps_lp_table, row.names=FALSE, quote=FALSE)

##Limit list to genes with an adjusted p value < 0.05
M_LPS.M_LP.sigGenes <- M_LPS.M_LP.topTab[M_LPS.M_LP.topTab$adj.P.Val < 0.05, ]
length(M_LPS.M_LP.sigGenes$ID)
##[1] 1473

##Filter out rows with less than 2-fold change (log2 fold change of > 1)
M_LPS.M_LP.sigGenesFold1 <- subset(M_LPS.M_LP.sigGenes, abs(logFC) > 1)
length(M_LPS.M_LP.sigGenesFold1$ID)
##[1] 467

##Filter out rows with less than 4-fold change (log2 fold change of > 2)
M_LPS.M_LP.sigGenesFold2 <- subset(M_LPS.M_LP.sigGenes, abs(logFC) > 2)
length(M_LPS.M_LP.sigGenesFold2$ID)
##[1] 125

##Make an MA plot
sel <- M_LPS.M_LP.topTab$adj.P.Val < 0.05
top <- M_LPS.M_LP.topTab
sub <- paste("No. of sig. genes: ", sum(sel), "/", length(sel))
cpm <- v$E

plot(rowMeans(cpm[top$ID,]), top$logFC, pch=16, cex=0.5, col="darkgrey",
     main="M_LPSvM_LP adjusted",
     ylab="log FC", xlab="Average Expression",
     sub=sub)
points(rowMeans(cpm[top$ID, ])[sel], top$logFC[sel], col="red", cex=0.5)
abline(h=c(-1, 0, 1), col="red")

dev.copy(pdf, "images/MAplot_M_LPSvM_LP_CDS_limmabatchcorrection_20171120rev.pdf",
         width=8, height=8)
dev.off()
dev.copy(png, "images/MAplot_M_LPSvM_LP_CDS_limmabatchcorrection_20171120rev.png",
         width=700, height=700)
dev.off()

M_LPS.M_LP.sigGenes <- M_LPS.M_LP.sigGenes[order(-M_LPS.M_LP.sigGenes$logFC), ]

##Annotate sigGenes list using Biomart
sigGenes <- M_LPS.M_LP.sigGenes
ids <- sigGenes$ID

##To see possibilities for attributes, use head(listAttributes(ensembl), n=20L)
desc <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol", "description", "gene_biotype"),
              filters="ensembl_gene_id",
              values=ids, mart=ensembl)
colnames(desc) <- c("ID", "Symbol", "Description", "Type")

##Remove commas from description
desc$Description <- gsub(",", "", desc$Description)

DEG <- merge(sigGenes, desc, by="ID", all=TRUE)
DEG <- subset(DEG, select=c(ID, Symbol, Description, logFC, adj.P.Val, AveExpr, Type), FS="/t")
DEG <- DEG[order(-DEG$logFC), ]

##Filter out Genes -1<FC<1
DEG <- subset(DEG, abs(DEG$logFC) > 1)

##Save DE genes
write.table(DEG, "csv/DEG_M_LPSvM_LP_CDS_limmabatchcorrection_20171120rev.txt",
            col.names=TRUE, row.names=FALSE, quote=FALSE)
write.csv(DEG, "csv/DEG_M_LPSvM_LP_CDS_limmabatchcorrection_20171120rev.csv",
          row.names=FALSE, quote=FALSE)
```

## M_LP v M_NS

```{r}
##eBayes finds an F-statistic from the set of t-statistics for that gene
M_LP.M_NS.contr.mat <- makeContrasts(M_LPvM_NS=(condM_LP - condM_NS), levels=v$design)
M_LP.M_NS.fit <- contrasts.fit(fit, M_LP.M_NS.contr.mat)
M_LP.M_NS.eb <- eBayes(M_LP.M_NS.fit)
M_LP.M_NS.topTab <- topTable(M_LP.M_NS.eb, coef="M_LPvM_NS", number=nrow(v$E))
M_LP.M_NS.topTab <- cbind(rownames(M_LP.M_NS.topTab), M_LP.M_NS.topTab)
colnames(M_LP.M_NS.topTab) <- c("ID", "logFC", "AveExpr", "t", "P.Value", "adj.P.Val", "B")
rownames(M_LP.M_NS.topTab) <- c(1:nrow(M_LP.M_NS.topTab))

lp_ns_table <- "csv/topTab_M_LPvM_NS_CDS_limmabatchcorrection_20171018.csv"
write.csv(M_LP.M_NS.topTab, file=lp_ns_table, row.names=FALSE, quote=FALSE)

##Limit list to genes with an adjusted p value < 0.05
M_LP.M_NS.sigGenes <- M_LP.M_NS.topTab[M_LP.M_NS.topTab$adj.P.Val < 0.05, ]
length(M_LP.M_NS.sigGenes$ID)
##[1] 4386

##Filter out rows with less than 2-fold change (log2 fold change of > 1)
M_LP.M_NS.sigGenesFold1 <- subset(M_LP.M_NS.sigGenes, abs(logFC) > 1)
length(M_LP.M_NS.sigGenesFold1$ID)
##[1] 1455

##Filter out rows with less than 4-fold change (log2 fold change of > 2)
M_LP.M_NS.sigGenesFold2 <- subset(M_LP.M_NS.sigGenes, abs(logFC) > 2)
length(M_LP.M_NS.sigGenesFold2$ID)
##[1] 519

##Make an MA plot
sel <- M_LP.M_NS.topTab$adj.P.Val < 0.05
top <- M_LP.M_NS.topTab
sub <- paste("No. of sig. genes: ", sum(sel), "/", length(sel))
cpm <- v$E

plot(rowMeans(cpm[top$ID,]), top$logFC, pch=16, cex=0.5, col="darkgrey",
     main="M_LPvM_NS adjusted",
     ylab="log FC", xlab="Average Expression",
     sub=sub)
points(rowMeans(cpm[top$ID,])[sel], top$logFC[sel], col="red", cex=0.5)
abline(h=c(-1, 0, 1), col="red")

dev.copy(pdf, "images/MAplot_M_LPvM_NS_CDS_limmabatchcorrection_20171018.pdf",
         width=8, height=8)
dev.off()
dev.copy(png, "images/MAplot_M_LPvM_NS_CDS_limmabatchcorrection_20171018.png",
         width=700, height=700)
dev.off()

M_LP.M_NS.sigGenes <- M_LP.M_NS.sigGenes[order(-M_LP.M_NS.sigGenes$logFC), ]

##Annotate sigGenes list using Biomart
sigGenes <- M_LP.M_NS.sigGenes
ids <- sigGenes$ID

##To see possibilities for attributes, use head(listAttributes(ensembl), n=20L)
desc <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol", "description", "gene_biotype"),
              filters="ensembl_gene_id",
              values=ids, mart=ensembl)
colnames(desc)=c("ID", "Symbol", "Description", "Type")

##Remove commas from description
desc$Description <- gsub(",", "", desc$Description)

DEG_LP4 <- merge(sigGenes, desc, by="ID", all=TRUE)
DEG_LP4 <- subset(DEG_LP4, select=c(ID, Symbol, Description, logFC, adj.P.Val, AveExpr, Type),
                  FS="/t")
DEG_LP4 <- DEG_LP4[order(-DEG_LP4$logFC), ]

##Filter out Genes -1<FC<1
DEG_LP4 <- subset(DEG_LP4, abs(DEG_LP4$logFC) > 1)

##Save DE genes
write.table(DEG_LP4, "csv/DEG_M_LPvM_NS_CDS_limmabatchcorrection_20171018.txt",
            col.names=TRUE, row.names=FALSE, quote=FALSE)
write.csv(DEG_LP4, "csv/DEG_M_LPvM_NS_CDS_limmabatchcorrection_20171018.csv",
          row.names=FALSE, quote=FALSE)
```

## M_LA v M_NS

```{r}
#eBayes finds an F-statistic from the set of t-statistics for that gene
M_LA.M_NS.contr.mat <- makeContrasts(M_LAvM_NS=(condM_LA - condM_NS), levels=v$design)
M_LA.M_NS.fit <- contrasts.fit(fit, M_LA.M_NS.contr.mat)
M_LA.M_NS.eb <- eBayes(M_LA.M_NS.fit)
M_LA.M_NS.topTab <- topTable(M_LA.M_NS.eb, coef="M_LAvM_NS", number=nrow(v$E))
M_LA.M_NS.topTab <- cbind(rownames(M_LA.M_NS.topTab), M_LA.M_NS.topTab)
colnames(M_LA.M_NS.topTab) <- c("ID", "logFC", "AveExpr", "t", "P.Value", "adj.P.Val", "B")
rownames(M_LA.M_NS.topTab) <- c(1:nrow(M_LA.M_NS.topTab))

la_ns_table <- "csv/topTab_M_LAvM_NS_CDS_limmabatchcorrection_20171120.csv"
write.csv(M_LA.M_NS.topTab, file=la_ns_table, row.names=FALSE, quote=FALSE)

##Limit list to genes with an adjusted p value < 0.05
M_LA.M_NS.sigGenes <- M_LA.M_NS.topTab[M_LA.M_NS.topTab$adj.P.Val < 0.05, ]
length(M_LA.M_NS.sigGenes$ID)
##[1] 617

##Filter out rows with less than 2-fold change (log2 fold change of > 1)
M_LA.M_NS.sigGenesFold1 <- subset(M_LA.M_NS.sigGenes, abs(logFC) > 1)
length(M_LA.M_NS.sigGenesFold1$ID)
##[1] 240

##Filter out rows with less than 4-fold change (log2 fold change of > 2)
M_LA.M_NS.sigGenesFold2 <- subset(M_LA.M_NS.sigGenes, abs(logFC) > 2)
length(M_LA.M_NS.sigGenesFold2$ID)
##[1] 69

##Make an MA plot
sel <- M_LA.M_NS.topTab$adj.P.Val < 0.05
top <- M_LA.M_NS.topTab
sub <- paste("No. of sig. genes: ", sum(sel), "/", length(sel))
cpm <- v$E

plot(rowMeans(cpm[top$ID,]), top$logFC, pch=16, cex=0.5, col="darkgrey",
     main="M_LAvM_NS adjusted",
     ylab="log FC", xlab="Average Expression",
     sub=sub)
points(rowMeans(cpm[top$ID, ])[sel], top$logFC[sel], col="red", cex=0.5)
abline(h=c(-1, 0, 1), col="red")

dev.copy(pdf, "images/MAplot_M_LAvM_NS_CDS_limmabatchcorrection_20171120.pdf", width=8, height=8)
dev.off()
dev.copy(png, "images/MAplot_M_LAvM_NS_CDS_limmabatchcorrection_20171120.png", width=700, height=700)
dev.off()

M_LA.M_NS.sigGenes <- M_LA.M_NS.sigGenes[order(-M_LA.M_NS.sigGenes$logFC), ]

##Annotate sigGenes list using Biomart
sigGenes <- M_LA.M_NS.sigGenes
ids <- sigGenes$ID

##To see possibilities for attributes, use head(listAttributes(ensembl), n=20L)
desc <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol", "description", "gene_biotype"),
              filters="ensembl_gene_id",
              values=ids, mart=ensembl)
colnames(desc) <- c("ID", "Symbol", "Description", "Type")

##Remove commas from description
desc$Description <- gsub(",", "", desc$Description)

DEG_LA4 <- merge(sigGenes, desc, by="ID", all=TRUE)
DEG_LA4 <- subset(DEG_LA4, select=c(ID, Symbol, Description, logFC, adj.P.Val, AveExpr, Type),
                  FS="/t")
DEG_LA4 <- DEG_LA4[order(-DEG_LA4$logFC), ]

##Filter out Genes -1<FC<1
DEG_LA4 <- subset(DEG_LA4, abs(DEG_LA4$logFC) > 1)

##Save DE genes
write.table(DEG_LA4, "csv/DEG_M_LAvM_NS_CDS_limmabatchcorrection_20171120.txt",
            col.names=TRUE, row.names=FALSE, quote=FALSE)
write.csv(DEG_LA4, "csv/DEG_M_LAvM_NS_CDS_limmabatchcorrection_20171120.csv",
          row.names=FALSE, quote=FALSE)
```

## GM_LPS v. GM_NS

```{r}
#eBayes finds an F-statistic from the set of t-statistics for that gene
GM_LPS.GM_NS.contr.mat <- makeContrasts(GM_LPSvGM_NS=(condGM_LPS-condGM_NS), levels=v$design)
GM_LPS.GM_NS.fit <- contrasts.fit(fit, GM_LPS.GM_NS.contr.mat)
GM_LPS.GM_NS.eb <- eBayes(GM_LPS.GM_NS.fit)
GM_LPS.GM_NS.topTab <- topTable(GM_LPS.GM_NS.eb, coef="GM_LPSvGM_NS", number=nrow(v$E))
GM_LPS.GM_NS.topTab <- cbind(rownames(GM_LPS.GM_NS.topTab), GM_LPS.GM_NS.topTab)
colnames(GM_LPS.GM_NS.topTab) <- c("ID", "logFC", "AveExpr", "t", "P.Value", "adj.P.Val", "B")
rownames(GM_LPS.GM_NS.topTab) <- c(1:nrow(GM_LPS.GM_NS.topTab))

gm_lps_ns_table <- "csv/topTab_GM_LPSvGM_NS_CDS_limmabatchcorrection_20171120.csv"
write.csv(GM_LPS.GM_NS.topTab, file= gm_lps_ns_table, row.names=FALSE, quote=FALSE)

##Limit list to genes with an adjusted p value < 0.05
GM_LPS.GM_NS.sigGenes <- GM_LPS.GM_NS.topTab[GM_LPS.GM_NS.topTab$adj.P.Val < 0.05, ]
length(GM_LPS.GM_NS.sigGenes$ID)
##[1] 302

##Filter out rows with less than 2-fold change (log2 fold change of > 1)
GM_LPS.GM_NS.sigGenesFold1 <- subset(GM_LPS.GM_NS.sigGenes, abs(logFC) > 1)
length(GM_LPS.GM_NS.sigGenesFold1$ID)
##[1] 197

##Filter out rows with less than 4-fold change (log2 fold change of > 2)
GM_LPS.GM_NS.sigGenesFold2 <- subset(GM_LPS.GM_NS.sigGenes, abs(logFC) > 2)
length(GM_LPS.GM_NS.sigGenesFold2$ID)
##[1] 101

##Make an MA plot
sel <- GM_LPS.GM_NS.topTab$adj.P.Val < 0.05
top <- GM_LPS.GM_NS.topTab
sub <- paste("No. of sig. genes: ", sum(sel), "/", length(sel))
cpm <- v$E

plot(rowMeans(cpm[top$ID,]), top$logFC, pch=16, cex=0.5, col="darkgrey",
     main="GM_LPSvGM_NS adjusted", ylab="log FC",
     xlab="Average Expression", sub=sub)
points(rowMeans(cpm[top$ID, ])[sel], top$logFC[sel], col="red", cex=0.5)
abline(h=c(-1, 0, 1), col="red")

dev.copy(pdf, "images/MAplot_GM_LPSvGM_NS_CDS_limmabatchcorrection_20171120.pdf",
         width=8, height=8)
dev.off()
dev.copy(png, "images/MAplot_GM_LPSvGM_NS_CDS_limmabatchcorrection_20171120.png",
         width=700, height=700)
dev.off()

GM_LPS.GM_NS.sigGenes <- GM_LPS.GM_NS.sigGenes[order(-GM_LPS.GM_NS.sigGenes$logFC), ]

##Annotate sigGenes list using Biomart
sigGenes <- GM_LPS.GM_NS.sigGenes
ids <- sigGenes$ID

##To see possibilities for attributes, use head(listAttributes(ensembl), n=20L)
desc <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol", "description", "gene_biotype"),
              filters="ensembl_gene_id",
              values=ids, mart=ensembl)
colnames(desc) <- c("ID", "Symbol", "Description", "Type")

##Remove commas from description
desc$Description <- gsub(",", "", desc$Description)

DEG <- merge(sigGenes, desc, by="ID", all=TRUE)
DEG <- subset(DEG, select=c(ID, Symbol, Description, logFC, adj.P.Val, AveExpr, Type), FS="/t")
DEG <- DEG[order(-DEG$logFC), ]

##Filter out Genes -1<FC<1
DEG <- subset(DEG, abs(DEG$logFC) > 1)

##Save DE genes
write.table(DEG, "csv/DEG_GM_LPSvGM_NS_CDS_limmabatchcorrection_20171120.txt",
            col.names=TRUE, row.names=FALSE, quote=FALSE)
write.csv(DEG, "csv/DEG_GM_LPSvGM_NS_CDS_limmabatchcorrection_20171120.csv",
          row.names=FALSE, quote=FALSE)
```

## GM_LPS v GM_LP

This contrast was written in the opposite order as the others, I think this is
the reason some of my plots keep looking backwards/upsidedown...  I am going to
comment out the original line and rewrite it so that it is identical in order to
what I found in: M_LP.M_LP.contr.mat.

```{r}
#eBayes finds an F-statistic from the set of t-statistics for that gene

## Here is the original line
## look here ----------------------------------------------->   v                     v
##GM_LPS.GM_LP.contr.mat <- makeContrasts(GM_LPSvGM_LP=((condGM_LPS-condGM_NS)-(condGM_LP-condGM_NS)),
##                                        levels=v$design)
##

## And here is my version:
## look here ----------------------------------------------->  v                     v
GM_LPS.GM_LP.contr.mat <- makeContrasts(
    GM_LPSvGM_LP=((condGM_LP - condGM_NS) - (condGM_LPS - condGM_NS)),
    levels=v$design)
## End of my change.

GM_LPS.GM_LP.fit <- contrasts.fit(fit, GM_LPS.GM_LP.contr.mat)
GM_LPS.GM_LP.eb <- eBayes(GM_LPS.GM_LP.fit)
GM_LPS.GM_LP.topTab <- topTable(GM_LPS.GM_LP.eb, coef="GM_LPSvGM_LP", number=nrow(v$E))
GM_LPS.GM_LP.topTab <- cbind(rownames(GM_LPS.GM_LP.topTab), GM_LPS.GM_LP.topTab)
colnames(GM_LPS.GM_LP.topTab) <- c("ID", "logFC", "AveExpr", "t", "P.Value", "adj.P.Val", "B")
rownames(GM_LPS.GM_LP.topTab) <- c(1:nrow(GM_LPS.GM_LP.topTab))

gm_lps_lp_table <- "csv/topTab_GM_LPSvGM_LP_CDS_limmabatchcorrection_20171023.csv"
write.csv(GM_LPS.GM_LP.topTab, file=gm_lps_lp_table, row.names=FALSE, quote=FALSE)

##Limit list to genes with an adjusted p value < 0.05
GM_LPS.GM_LP.sigGenes <- GM_LPS.GM_LP.topTab[GM_LPS.GM_LP.topTab$adj.P.Val < 0.05, ]
length(GM_LPS.GM_LP.sigGenes$ID)
##[1] 194

##Filter out rows with less than 2-fold change (log2 fold change of > 1)
GM_LPS.GM_LP.sigGenesFold1 <- subset(GM_LPS.GM_LP.sigGenes, abs(logFC) > 1)
length(GM_LPS.GM_LP.sigGenesFold1$ID)
##[1] 114

##Filter out rows with less than 4-fold change (log2 fold change of > 2)
GM_LPS.GM_LP.sigGenesFold2 <- subset(GM_LPS.GM_LP.sigGenes, abs(logFC) > 2)
length(GM_LPS.GM_LP.sigGenesFold2$ID)
##[1] 27

##Make an MA plot
sel <- GM_LPS.GM_LP.topTab$adj.P.Val < 0.05
top <- GM_LPS.GM_LP.topTab
sub <- paste("No. of sig. genes: ", sum(sel), "/", length(sel))
cpm <- v$E

plot(rowMeans(cpm[top$ID,]), top$logFC, pch=16, cex=0.5, col="darkgrey",
     main="GM_LPSvGM_LP adjusted",
     ylab="log FC", xlab="Average Expression",
     sub=sub)
points(rowMeans(cpm[top$ID, ])[sel], top$logFC[sel], col="red", cex=0.5)
abline(h=c(-1, 0, 1), col="red")

dev.copy(pdf, "images/MAplot_GM_LPSvGM_LP_CDS_limmabatchcorrection_20171120.pdf",
         width=8, height=8)
dev.off()
dev.copy(png, "images/MAplot_GM_LPSvGM_LP_CDS_limmabatchcorrection_20171120.png",
         width=700, height=700)
dev.off()

GM_LPS.GM_LP.sigGenes <- GM_LPS.GM_LP.sigGenes[order(-GM_LPS.GM_LP.sigGenes$logFC), ]

## Annotate sigGenes list using Biomart
sigGenes <- GM_LPS.GM_LP.sigGenes
ids <- sigGenes$ID

## To see possibilities for attributes, use head(listAttributes(ensembl), n=20L)
desc <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol", "description", "gene_biotype"),
              filters="ensembl_gene_id", values=ids, mart=ensembl)
colnames(desc) <- c("ID", "Symbol", "Description", "Type")

##Remove commas from description
desc$Description <- gsub(",", "", desc$Description)

DEG <- merge(sigGenes, desc, by="ID", all=TRUE)
DEG <- subset(DEG, select=c(ID, Symbol, Description, logFC, adj.P.Val, AveExpr, Type), FS="/t")
DEG <- DEG[order(-DEG$logFC), ]

##Filter out Genes -1<FC<1
DEG <- subset(DEG, abs(DEG$logFC) > 1)

##Save DE genes
write.table(DEG, "csv/DEG_GM_LPSvGM_LP_CDS_limmabatchcorrection_20171120.txt",
            col.names=TRUE, row.names=FALSE, quote=FALSE)
write.csv(DEG, "csv/DEG_GM_LPSvGM_LP_CDS_limmabatchcorrection_20171120.csv",
          row.names=FALSE, quote=FALSE)
```

## GM_LPS v GM_LA

I think I observe the same flipping here.

```{r}
## eBayes finds an F-statistic from the set of t-statistics for that gene
## GM_LPS.GM_LA.contr.mat <- makeContrasts(GM_LPSvGM_LA=((condGM_LPS-condGM_NS)-(condGM_LA-condGM_NS)),
##                                         levels=v$design)
GM_LPS.GM_LA.contr.mat <- makeContrasts(
    GM_LPSvGM_LA=((condGM_LA - condGM_NS) - (condGM_LPS - condGM_NS)),
    levels=v$design)
## End of my change
GM_LPS.GM_LA.fit <- contrasts.fit(fit, GM_LPS.GM_LA.contr.mat)
GM_LPS.GM_LA.eb <- eBayes(GM_LPS.GM_LA.fit)
GM_LPS.GM_LA.topTab <- topTable(GM_LPS.GM_LA.eb, coef="GM_LPSvGM_LA", number=nrow(v$E))
GM_LPS.GM_LA.topTab <- cbind(rownames(GM_LPS.GM_LA.topTab), GM_LPS.GM_LA.topTab)
colnames(GM_LPS.GM_LA.topTab) <- c("ID", "logFC", "AveExpr", "t", "P.Value", "adj.P.Val", "B")
rownames(GM_LPS.GM_LA.topTab) <- c(1:nrow(GM_LPS.GM_LA.topTab))

gm_lps_la_table <- "csv/topTab_GM_LPSvGM_LA_CDS_limmabatchcorrection_20171023.csv"
write.csv(GM_LPS.GM_LA.topTab, file=gm_lps_la_table, row.names=FALSE, quote=FALSE)

## Limit list to genes with an adjusted p value < 0.05
GM_LPS.GM_LA.sigGenes <- GM_LPS.GM_LA.topTab[GM_LPS.GM_LA.topTab$adj.P.Val < 0.05, ]
length(GM_LPS.GM_LA.sigGenes$ID)
## [1] 17

## Filter out rows with less than 2-fold change (log2 fold change of > 1)
GM_LPS.GM_LA.sigGenesFold1 <- subset(GM_LPS.GM_LA.sigGenes, abs(logFC) > 1)
length(GM_LPS.GM_LA.sigGenesFold1$ID)
## [1] 15

## Filter out rows with less than 4-fold change (log2 fold change of > 2)
GM_LPS.GM_LA.sigGenesFold2 <- subset(GM_LPS.GM_LA.sigGenes, abs(logFC) > 2)
length(GM_LPS.GM_LA.sigGenesFold2$ID)
## [1] 7

## Make an MA plot
sel <- GM_LPS.GM_LA.topTab$adj.P.Val < 0.05
top <- GM_LPS.GM_LA.topTab
sub <- paste("No. of sig. genes: ", sum(sel),"/", length(sel))
cpm <- v$E

plot(rowMeans(cpm[top$ID,]), top$logFC, pch=16, cex=0.5, col="darkgrey",
     main="GM_LPSvGM_LA adjusted", ylab="log FC", xlab="Average Expression", sub=sub)
points(rowMeans(cpm[top$ID,])[sel], top$logFC[sel], col="red", cex=0.5)
abline(h=c(-1, 0, 1), col="red")

dev.copy(pdf, "images/MAplot_GM_LPSvGM_LA_CDS_limmabatchcorrection_20171120.pdf",
         width=8, height=8)
dev.off()
dev.copy(png, "images/MAplot_GM_LPSvGM_LA_CDS_limmabatchcorrection_20171120.png",
         width=700, height=700)
dev.off()

GM_LPS.GM_LA.sigGenes <- GM_LPS.GM_LA.sigGenes[order(-GM_LPS.GM_LA.sigGenes$logFC), ]

## Annotate sigGenes list using Biomart
sigGenes <- GM_LPS.GM_LA.sigGenes
ids <- sigGenes$ID

## To see possibilities for attributes, use head(listAttributes(ensembl), n=20L)
desc <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol", "description", "gene_biotype"),
              filters="ensembl_gene_id", values=ids, mart=ensembl)
colnames(desc) <- c("ID", "Symbol", "Description", "Type")

## Remove commas from description
desc$Description <- gsub(",", "", desc$Description)

DEG <- merge(sigGenes, desc, by="ID", all=TRUE)
DEG <- subset(DEG, select=c(ID, Symbol, Description, logFC, adj.P.Val, AveExpr, Type), FS="/t")
DEG <- DEG[order(-DEG$logFC), ]

## Filter out Genes -1<FC<1
DEG <- subset(DEG, abs(DEG$logFC) > 1)

## Save DE genes
write.table(DEG, "csv/DEG_GM_LPSvGM_LA_CDS_limmabatchcorrection_20171120.txt",
            col.names=TRUE, row.names=FALSE, quote=FALSE)
write.csv(DEG, "csv/DEG_GM_LPSvGM_LA_CDS_limmabatchcorrection_20171120.csv",
          row.names=FALSE, quote=FALSE)
```

## Barplots of Fold-Changes, e.g. Figure 3C/D

It appears to me that these figures are created by merging the M/GM tables for
the LP-LPS and LA-LPS contrasts, taking the top ~20 for the M table, and
plotting the log2FC on the linear scale.  I believe I can trivially modify this
to add the logFC / t-stat.  The caveat will be that doing this gives me a log2
error bar, not linear...  My inclination therefore is to just plot the logFC
rather than convert it to linear, but whatever.

### Figure 3C

Najib and Dave want some significance test between the M/GM bars for these 20
genes.  As I think about it, I think the only valid way to do this is to
calculate the logFC of GM(PG/LP) / M(PG/LP).  Then let limma calculate the t
statistic and p-value using its voom() modified values.

```{r my_fig3c_stats}
GM_PG.M_PG.contr.mat <- makeContrasts(
    GM_PGvM_PG=((condGM_LP - condGM_LPS) - (condM_LP - condM_LPS)),
    levels=v$design)
GM_PG.M_PG.fit <- contrasts.fit(fit, GM_PG.M_PG.contr.mat)
GM_PG.M_PG.eb <- eBayes(GM_PG.M_PG.fit)
GM_PG.M_PG.topTab <- topTable(GM_PG.M_PG.eb, coef="GM_PGvM_PG", number=nrow(v$E))

desc <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol", "description", "gene_biotype"),
              filters="ensembl_gene_id", values=rownames(GM_PG.M_PG.topTab), mart=ensembl)
colnames(desc) <- c("ID", "Symbol", "Description", "Type")
gene_symbols <- desc[, c("ID", "Symbol")]

GM_PG.M_PG.topTab <- merge(gene_symbols, GM_PG.M_PG.topTab, by.x="ID", by.y="row.names")

GM_LA.M_LA.contr.mat <- makeContrasts(
    GM_LAvM_LA=((condGM_LA - condGM_LPS) - (condM_LA - condM_LPS)),
    levels=v$design)
GM_LA.M_LA.fit <- contrasts.fit(fit, GM_LA.M_LA.contr.mat)
GM_LA.M_LA.eb <- eBayes(GM_LA.M_LA.fit)
GM_LA.M_LA.topTab <- topTable(GM_LA.M_LA.eb, coef="GM_LAvM_LA", number=nrow(v$E))

GM_LA.M_LA.topTab <- merge(gene_symbols, GM_LA.M_LA.topTab, by.x="ID", by.y="row.names")
```

### Figure 3C top-left

Kajal's variables to create this are...

1.  M_LPS.M_LP.topTab
2.  GM_LPS.GM_LP.topTab

These were passed to biomart to get the gene names rather than ensembl IDs.
I am going to be lazy and just copy/paste Kajal's code for these tasks.

Having done the top-left piece of this, it seems to me that if you are going to
take the top 20 genes and exclude based on the M-CSF adjusted p-value, perhaps
you should also exclude based on the GM-CSF adjusted p-value, but the way this
was done, only the M is used.  In its current state, I am only using the M as
per the figures in their current state.

Note, that if you want error bars from the standard error, then it is waaaay
easier to stay on the log2 scale rather than convert back to linear because the
math for converting the standard error back to linear is weird.  If the table
had a few more parameters in it, I could do it without struggling, but it doesn't.

```{r fig3c_topleft}
wanted_columns <- c("ID", "logFC.x", "logFC.y", "t.x", "t.y", "Symbol")
renamed_columns <- c("ID", "m_logfc", "gm_logfc", "m_t", "gm_t", "symbol")

fig3c_lp_df <- merge(M_LPS.M_LP.topTab, GM_LPS.GM_LP.topTab, by="ID")
desc <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol", "description", "gene_biotype"),
              filters="ensembl_gene_id", values=fig3c_lp_df[["ID"]], mart=ensembl)
colnames(desc) <- c("ID", "Symbol", "Description", "Type")
fig3c_lp_df <- merge(fig3c_lp_df, desc, by="ID", all.x=TRUE)

tl_order_idx <- order(fig3c_lp_df[["logFC.x"]], decreasing=TRUE)
fig3c_tl_df <- fig3c_lp_df[tl_order_idx, ]
sig_idx <- fig3c_tl_df[["adj.P.Val.x"]] <= 0.05
fig3c_tl_df <- head(fig3c_tl_df[sig_idx, ], n=25)
fig3c_tl_df <- fig3c_tl_df[, wanted_columns]
colnames(fig3c_tl_df) <- renamed_columns
rownames(fig3c_tl_df) <- fig3c_tl_df[["ID"]]
fig3c_tl_df[["ID"]] <- NULL
fig3c_tl_df[["m_lfcerr"]] <- fig3c_tl_df[["m_logfc"]] / fig3c_tl_df[["m_t"]]
fig3c_tl_df[["gm_lfcerr"]] <- fig3c_tl_df[["gm_logfc"]] / fig3c_tl_df[["gm_t"]]

fig3c_tl_m <- fig3c_tl_df[, c("m_logfc", "symbol", "m_lfcerr")]
fig3c_tl_m[["type"]] <- "m"
colnames(fig3c_tl_m) <- c("logfc", "symbol", "lfcerr", "type")
fig3c_tl_gm <- fig3c_tl_df[, c("gm_logfc", "symbol", "gm_lfcerr")]
fig3c_tl_gm[["type"]] <- "gm"
colnames(fig3c_tl_gm) <- c("logfc", "symbol", "lfcerr", "type")
melted_fig3c_tl <- rbind(fig3c_tl_m, fig3c_tl_gm)
melted_fig3c_tl[["symbol"]] <- factor(melted_fig3c_tl[["symbol"]],
                                      levels=fig3c_tl_m[["symbol"]])
melted_fig3c_tl[["type"]] <- factor(melted_fig3c_tl[["type"]],
                                    levels=c("m", "gm"))

p <- ggplot(data=melted_fig3c_tl, aes(x=symbol, y=logfc, fill=type)) +
  geom_bar(stat="identity", color="black", position=position_dodge())+
  scale_fill_manual(values=c("cornflowerblue", "darkgrey")) +
  geom_errorbar(aes(ymin=logfc - (lfcerr / 2), ymax=logfc + (lfcerr / 2)), width=.2,
                position=position_dodge(0.9)) +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=0.5))
p

sig_idx <- GM_PG.M_PG.topTab[["Symbol"]] %in% fig3c_tl_gm[["symbol"]]
sig_values <- GM_PG.M_PG.topTab[sig_idx, c("Symbol", "P.Value", "adj.P.Val")]
rownames(sig_values) <- sig_values[["Symbol"]]
sig_values[levels(melted_fig3c_tl[["symbol"]]), ]
```

### Figure 3C bottom-left

```{r fig3C_bottomleft}
bl_order_idx <- order(fig3c_lp_df[["logFC.x"]], decreasing=FALSE)
fig3c_bl_df <- fig3c_lp_df[bl_order_idx, ]
sig_idx <- fig3c_bl_df[["adj.P.Val.x"]] <= 0.05
fig3c_bl_df <- head(fig3c_bl_df[sig_idx, ], n=20)
fig3c_bl_df <- fig3c_bl_df[, wanted_columns]
colnames(fig3c_bl_df) <- renamed_columns
fig3c_bl_df[["ID"]] <- NULL
fig3c_bl_df[["m_lfcerr"]] <- fig3c_bl_df[["m_logfc"]] / fig3c_bl_df[["m_t"]]
fig3c_bl_df[["gm_lfcerr"]] <- fig3c_bl_df[["gm_logfc"]] / fig3c_bl_df[["gm_t"]]

fig3c_bl_m <- fig3c_bl_df[, c("m_logfc", "symbol", "m_lfcerr")]
fig3c_bl_m[["type"]] <- "m"
colnames(fig3c_bl_m) <- c("logfc", "symbol", "lfcerr", "type")
fig3c_bl_gm <- fig3c_bl_df[, c("gm_logfc", "symbol", "gm_lfcerr")]
fig3c_bl_gm[["type"]] <- "gm"
colnames(fig3c_bl_gm) <- c("logfc", "symbol", "lfcerr", "type")
melted_fig3c_bl <- rbind(fig3c_bl_m, fig3c_bl_gm)
melted_fig3c_bl[["symbol"]] <- factor(melted_fig3c_bl[["symbol"]],
                                      levels=fig3c_bl_m[["symbol"]])
melted_fig3c_bl[["type"]] <- factor(melted_fig3c_bl[["type"]],
                                    levels=c("m", "gm"))

p <- ggplot(data=melted_fig3c_bl, aes(x=symbol, y=logfc, fill=type)) +
  geom_bar(stat="identity", color="black", position=position_dodge())+
  scale_fill_manual(values=c("cornflowerblue", "darkgrey")) +
  geom_errorbar(aes(ymin=logfc-lfcerr, ymax=logfc+lfcerr), width=.2,
                position=position_dodge(0.9)) +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=0.5))
p

sig_idx <- GM_PG.M_PG.topTab[["Symbol"]] %in% fig3c_bl_gm[["symbol"]]
sig_values <- GM_PG.M_PG.topTab[sig_idx, c("Symbol", "P.Value", "adj.P.Val")]
rownames(sig_values) <- sig_values[["Symbol"]]
sig_values[levels(melted_fig3c_bl[["symbol"]]), ]
```

### Figure 3C top-right

The right side of the plot is LA rather than LP, otherwise this is the same.

```{r fig3c_topright}
wanted_columns <- c("ID", "logFC.x", "logFC.y", "t.x", "t.y", "Symbol")
renamed_columns <- c("ID", "m_logfc", "gm_logfc", "m_t", "gm_t", "symbol")

fig3c_la_df <- merge(M_LPS.M_LA.topTab, GM_LPS.GM_LA.topTab, by="ID")
desc <- getBM(attributes=c("ensembl_gene_id", "hgnc_symbol", "description", "gene_biotype"),
              filters="ensembl_gene_id", values=fig3c_la_df[["ID"]], mart=ensembl)
colnames(desc) <- c("ID", "Symbol", "Description", "Type")
fig3c_la_df <- merge(fig3c_la_df, desc, by="ID", all.x=TRUE)

tr_order_idx <- order(fig3c_la_df[["logFC.x"]], decreasing=TRUE)
fig3c_tr_df <- fig3c_la_df[tr_order_idx, ]
sig_idx <- fig3c_tr_df[["adj.P.Val.x"]] <= 0.05
fig3c_tr_df <- head(fig3c_tr_df[sig_idx, ], n=25)
fig3c_tr_df <- fig3c_tr_df[, wanted_columns]
colnames(fig3c_tr_df) <- renamed_columns
rownames(fig3c_tr_df) <- fig3c_tr_df[["ID"]]
fig3c_tr_df[["ID"]] <- NULL
fig3c_tr_df[["m_lfcerr"]] <- fig3c_tr_df[["m_logfc"]] / fig3c_tr_df[["m_t"]]
fig3c_tr_df[["gm_lfcerr"]] <- fig3c_tr_df[["gm_logfc"]] / fig3c_tr_df[["gm_t"]]

fig3c_tr_m <- fig3c_tr_df[, c("m_logfc", "symbol", "m_lfcerr")]
fig3c_tr_m[["type"]] <- "m"
colnames(fig3c_tr_m) <- c("logfc", "symbol", "lfcerr", "type")
fig3c_tr_gm <- fig3c_tr_df[, c("gm_logfc", "symbol", "gm_lfcerr")]
fig3c_tr_gm[["type"]] <- "gm"
colnames(fig3c_tr_gm) <- c("logfc", "symbol", "lfcerr", "type")
melted_fig3c_tr <- rbind(fig3c_tr_m, fig3c_tr_gm)
melted_fig3c_tr[["symbol"]] <- factor(melted_fig3c_tr[["symbol"]],
                                      levels=fig3c_tr_m[["symbol"]])
melted_fig3c_tr[["type"]] <- factor(melted_fig3c_tr[["type"]],
                                    levels=c("m", "gm"))

p <- ggplot(data=melted_fig3c_tr, aes(x=symbol, y=logfc, fill=type)) +
  geom_bar(stat="identity", color="black", position=position_dodge())+
  scale_fill_manual(values=c("cornflowerblue", "darkgrey")) +
  geom_errorbar(aes(ymin=logfc-lfcerr, ymax=logfc+lfcerr), width=.2,
                position=position_dodge(0.9)) +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=0.5))
p

sig_idx <- GM_LA.M_LA.topTab[["Symbol"]] %in% fig3c_tr_gm[["symbol"]]
sig_values <- GM_LA.M_LA.topTab[sig_idx, c("Symbol", "P.Value", "adj.P.Val")]
rownames(sig_values) <- sig_values[["Symbol"]]
sig_values[levels(melted_fig3c_tr[["symbol"]]), ]
```

### Figure 3C bottom-right

```{r fig3C_bottomright}
br_order_idx <- order(fig3c_la_df[["logFC.x"]], decreasing=FALSE)
fig3c_br_df <- fig3c_la_df[br_order_idx, ]
sig_idx <- fig3c_br_df[["adj.P.Val.x"]] <= 0.05
fig3c_br_df <- head(fig3c_br_df[sig_idx, ], n=20)
fig3c_br_df <- fig3c_br_df[, wanted_columns]
colnames(fig3c_br_df) <- renamed_columns
fig3c_br_df[["ID"]] <- NULL
fig3c_br_df[["m_lfcerr"]] <- fig3c_br_df[["m_logfc"]] / fig3c_br_df[["m_t"]]
fig3c_br_df[["gm_lfcerr"]] <- fig3c_br_df[["gm_logfc"]] / fig3c_br_df[["gm_t"]]

fig3c_br_m <- fig3c_br_df[, c("m_logfc", "symbol", "m_lfcerr")]
fig3c_br_m[["type"]] <- "m"
colnames(fig3c_br_m) <- c("logfc", "symbol", "lfcerr", "type")
fig3c_br_gm <- fig3c_br_df[, c("gm_logfc", "symbol", "gm_lfcerr")]
fig3c_br_gm[["type"]] <- "gm"
colnames(fig3c_br_gm) <- c("logfc", "symbol", "lfcerr", "type")
melted_fig3c_br <- rbind(fig3c_br_m, fig3c_br_gm)
melted_fig3c_br[["symbol"]] <- factor(melted_fig3c_br[["symbol"]],
                                      levels=fig3c_br_m[["symbol"]])
melted_fig3c_br[["type"]] <- factor(melted_fig3c_br[["type"]],
                                    levels=c("m", "gm"))

p <- ggplot(data=melted_fig3c_br, aes(x=symbol, y=logfc, fill=type)) +
  geom_bar(stat="identity", color="black", position=position_dodge())+
  scale_fill_manual(values=c("cornflowerblue", "darkgrey")) +
  geom_errorbar(aes(ymin=logfc-lfcerr, ymax=logfc+lfcerr), width=.2,
                position=position_dodge(0.9)) +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=0.5))
p

sig_idx <- GM_LA.M_LA.topTab[["Symbol"]] %in% fig3c_br_gm[["symbol"]]
sig_values <- GM_LA.M_LA.topTab[sig_idx, c("Symbol", "P.Value", "adj.P.Val")]
rownames(sig_values) <- sig_values[["Symbol"]]
sig_values[levels(melted_fig3c_br[["symbol"]]), ]
```

### Figure 2A

Kajal responded to my query about this figure, and I am a little embarrassed to
say that I should have seen what it is.  It seems to me that it is a bit
redundant though with 3c, as it is precisely a subset of that data, though
slightly reordered.

#### Figure 2A Left

```{r fig2a1}
fig2a_df <- M_LPS.M_LP.sigGenes
order_idx <- order(fig2a_df[["logFC"]], decreasing=FALSE)
fig2a_df <- fig2a_df[order_idx, ]
top <- head(fig2a_df, n=10)
bot <- tail(fig2a_df, n=10)
plotted <- rbind(top, bot)
rownames(plotted) <- plotted[["ID"]]
desc_idx <- desc[["ID"]] %in% plotted[["ID"]]
this_desc <- desc[desc_idx, ]
current <- rownames(plotted)
plotted <- merge(this_desc, plotted, by.x="ID", by.y="ID", all=TRUE)
rownames(plotted) <- make.names(plotted[["ID"]], unique=TRUE)
plotted <- plotted[current, ]
plotted[["lfcerr"]] <- plotted[["logFC"]] / plotted[["t"]]
plotted[["Symbol"]] <- factor(plotted[["Symbol"]], levels=plotted[["Symbol"]])

p <- ggplot(data=plotted, aes(x=Symbol, y=logFC)) +
  geom_bar(stat="identity", fill="#89A8DD", color="black", position=position_dodge())+
  geom_errorbar(aes(ymin=logFC - (lfcerr / 2), ymax=logFC + (lfcerr / 2)), width=0.2,
                position=position_dodge(0.9)) +
  coord_flip() +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=0.5))
p
```

#### Figure 2A Right

```{r fig2a}
fig2a_df <- M_LPS.M_LA.sigGenes
order_idx <- order(fig2a_df[["logFC"]], decreasing=FALSE)
fig2a_df <- fig2a_df[order_idx, ]
top <- head(fig2a_df, n=10)
bot <- tail(fig2a_df, n=10)
plotted <- rbind(top, bot)
rownames(plotted) <- plotted[["ID"]]
desc_idx <- desc[["ID"]] %in% plotted[["ID"]]
this_desc <- desc[desc_idx, ]
current <- rownames(plotted)
plotted <- merge(this_desc, plotted, by.x="ID", by.y="ID", all=TRUE)
rownames(plotted) <- make.names(plotted[["ID"]], unique=TRUE)
plotted <- plotted[current, ]
plotted[["lfcerr"]] <- plotted[["logFC"]] / plotted[["t"]]
plotted[["Symbol"]] <- factor(plotted[["Symbol"]], levels=plotted[["Symbol"]])

p <- ggplot(data=plotted, aes(x=Symbol, y=logFC)) +
  geom_bar(stat="identity", fill="#7BD19C", color="black", position=position_dodge())+
  geom_errorbar(aes(ymin=logFC - (lfcerr / 2), ymax=logFC + (lfcerr / 2)), width=0.2,
                position=position_dodge(0.9)) +
  coord_flip() +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=0.5))
p
```

### Figure 2D

I created a table of shared DEG between L+PGE2/LPS and L+Ado/LPS (table attached
below) by merging the 2 DE tables based on common gene symbol.  In this table,
logFC.x corresponds to L+Ado/LPS and logFC.y corresponds to L+PGE2/LPS.  I
averaged the fold changes between logFC.x and logFC.y in order to determine the
rank order of the genes in the figure but plotted the fold changes
individually.  Instead of plotting log2FC I manually converted those values to
just FC. I dont know how to calculate standard error if I am plotting just FC.
It would be nice if this could all be done in R.

```{r fig2dt}
fig2d_df <- merge(M_LPS.M_LP.sigGenes, M_LPS.M_LA.sigGenes, by="ID")
fig2d_df <- merge(fig2d_df, desc, by="ID")
fig2d_df[["avg"]] <- (fig2d_df[["logFC.x"]] + fig2d_df[["logFC.y"]]) / 2

order_idx <- order(fig2d_df[["avg"]], decreasing=TRUE)
fig2dt_df <- head(fig2d_df[order_idx, ], n=22)
fig2dt_df[["lp_lfcerr"]] <- fig2dt_df[["logFC.x"]] / fig2dt_df[["t.x"]]
fig2dt_df[["la_lfcerr"]] <- fig2dt_df[["logFC.y"]] / fig2dt_df[["t.y"]]

fig2dt_lp <- fig2dt_df[, c("Symbol", "logFC.x", "lp_lfcerr")]
colnames(fig2dt_lp) <- c("symbol", "logfc", "lfcerr")
fig2dt_lp[["type"]] <- "LPS+PGE"
fig2dt_la <- fig2dt_df[, c("Symbol", "logFC.y", "la_lfcerr")]
colnames(fig2dt_la) <- c("symbol", "logfc", "lfcerr")
fig2dt_la[["type"]] <- "LPS+Ado"
melted_fig2dt <- rbind(fig2dt_la, fig2dt_lp)
melted_fig2dt[["symbol"]] <- factor(melted_fig2dt[["symbol"]],
                                    levels=unique(melted_fig2dt[["symbol"]]))
melted_fig2dt[["type"]] <- factor(melted_fig2dt[["type"]], levels=c("LPS+Ado", "LPS+PGE"))

p <- ggplot(data=melted_fig2dt, aes(x=symbol, y=logfc, fill=type)) +
  geom_bar(stat="identity", color="black", position=position_dodge())+
  scale_fill_manual(values=c("#7BD19C", "#89A8DD")) +
  geom_errorbar(aes(ymin=logfc - (lfcerr / 2), ymax=logfc + (lfcerr / 2)), width=0.2,
                position=position_dodge(0.9)) +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=0.5))
p
```

```{r fig2db}
fig2d_df <- merge(M_LPS.M_LP.sigGenes, M_LPS.M_LA.sigGenes, by="ID")
fig2d_df <- merge(fig2d_df, desc, by="ID")
fig2d_df[["avg"]] <- (fig2d_df[["logFC.x"]] + fig2d_df[["logFC.y"]]) / 2

order_idx <- order(fig2d_df[["avg"]], decreasing=FALSE)
fig2db_df <- head(fig2d_df[order_idx, ], n=22)
fig2db_df[["lp_lfcerr"]] <- fig2db_df[["logFC.x"]] / fig2db_df[["t.x"]]
fig2db_df[["la_lfcerr"]] <- fig2db_df[["logFC.y"]] / fig2db_df[["t.y"]]

fig2db_lp <- fig2db_df[, c("Symbol", "logFC.x", "lp_lfcerr")]
colnames(fig2db_lp) <- c("symbol", "logfc", "lfcerr")
fig2db_lp[["type"]] <- "LPS+PGE"
fig2db_la <- fig2db_df[, c("Symbol", "logFC.y", "la_lfcerr")]
colnames(fig2db_la) <- c("symbol", "logfc", "lfcerr")
fig2db_la[["type"]] <- "LPS+Ado"
melted_fig2db <- rbind(fig2db_la, fig2db_lp)
melted_fig2db[["symbol"]] <- factor(melted_fig2db[["symbol"]],
                                    levels=unique(melted_fig2db[["symbol"]]))
melted_fig2db[["type"]] <- factor(melted_fig2db[["type"]], levels=c("LPS+Ado", "LPS+PGE"))

p <- ggplot(data=melted_fig2db, aes(x=symbol, y=logfc, fill=type)) +
  geom_bar(stat="identity", color="black", position=position_dodge())+
  scale_fill_manual(values=c("#7BD19C", "#89A8DD")) +
  geom_errorbar(aes(ymin=logfc - (lfcerr / 2), ymax=logfc + (lfcerr / 2)), width=0.2,
                position=position_dodge(0.9)) +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=0.5))
p
```

## Venn Diagram of shared DEG between M-CSF LPS v LA v LP

Upregulated genes compared to NS

```{r}
v_data <- list(DEG_LA4[DEG_LA4$logFC > 0, ]$ID,
               DEG_LP4[DEG_LP4$logFC > 0, ]$ID,
               DEG_LPS[DEG_LPS$logFC > 0, ]$ID)
v_data <- Venn(v_data, SetNames=c("LPS+Ado", "LPS+PGE2", "LPS"), numberOfSets=3)
Vennerable::plot(v_data, doWeights=FALSE)
## dev.copy(png, "VennDiagram Upregulated LPS LA LP", width=700, height=700)
## dev.off();
```

Downregulated genes compared to NS

```{r}
v_data <- list(DEG_LA4[DEG_LA4$logFC < 0, ]$ID,
               DEG_LP4[DEG_LP4$logFC < 0, ]$ID,
               DEG_LPS[DEG_LPS$logFC < 0, ]$ID)
v_data <- Venn(v_data, SetNames=c("LPS+Ado", "LPS+PGE2", "LPS"), numberOfSets=3)
Vennerable::plot(v_data, doWeights=FALSE)
## dev.copy(png, "VennDiagram Downregulated LPS LA LP", width=700, height=700)
## dev.off();
```

## Volcano Plots

M_LPS v M_LA

```{r previous_volcano}
res <- read.csv(lps_la_table)
with(res, plot(logFC, -log10(P.Value), cex=0.8, pch=20,
               main="Volcano plot M_LPS v M_LA", xlim=c(-7, 7)))
## Add colored points: red if padj<0.05, orange of log2FC>1, green if both)
with(subset(res, adj.P.Val < 0.05 ),
     points(logFC, -log10(P.Value), cex=0.8, pch=20, col="violetred2"))
with(subset(res, abs(logFC) > 1),
     points(logFC, -log10(P.Value), cex=0.8, pch=20, col="orange"))
with(subset(res, adj.P.Val < 0.05 & abs(logFC) > 1),
     points(logFC, -log10(P.Value), cex=0.8, pch=20, col="turquoise3"))
dev.copy(png, "images/VolcanoPlot_M_LPS_v_M_LA.png", width=700, height=700)
dev.off()
```

```{r print_numbers_downla}
sig_up_idx <- res[, "adj.P.Val"] < 0.05 & res[, "logFC"] > 1
sum(sig_up_idx)
sig_up_ids <- res[sig_up_idx, "ID"]
m_lps_m_la_up_ids <- sig_up_ids

sig_down_idx <- res[, "adj.P.Val"] < 0.05 & res[, "logFC"] < -1
sum(sig_down_idx)
sig_down_ids <- res[sig_down_idx, "ID"]
m_lps_m_la_down_ids <- sig_down_ids
```

# M_LPS v M_LP

```{r uneval_volcano}
res <- read.csv("csv/topTab_M_LPSvM_LP_CDS_limmabatchcorrection_20171120rev.csv", header=TRUE)
head(res)
with(res, plot(logFC, -log10(P.Value), cex=0.8, pch=20,
               main="Volcano plot M_LPS v M_LP", xlim=c(-7, 7)))

## Add colored points: red if padj<0.05, orange of log2FC>1, green if both)
with(subset(res, adj.P.Val < 0.05 ),
     points(logFC, -log10(P.Value), cex=0.8, pch=20, col="violetred2"))
with(subset(res, abs(logFC) > 1),
     points(logFC, -log10(P.Value), cex=0.8, pch=20, col="orange"))
with(subset(res, adj.P.Val < 0.05 & abs(logFC) > 1),
     points(logFC, -log10(P.Value), cex=0.8, pch=20, col="turquoise3"))
dev.copy(png, "images/VolcanoPlot_M_LPS_v_M_LP.png", width=700, height=700)
dev.off()
```

## Rich Factor Graphs

## GO terms shared LA LP

```{r}
GOSharedpathways <- read.table("csv/GO MF FDR .05 TOP 5 only-2.txt", header=TRUE, sep='\t')
GOSharedpathways <- GOSharedpathways[order(GOSharedpathways$p.Value, decreasing=TRUE), ]

colorRamp <- colorRampPalette(c("turquoise1", "mediumblue"))(50)
par(mar=c(5, 18, 3, 8))
plot(GOSharedpathways$Rich.Factor, c(1:5), axes=FALSE,
     panel.first=abline(h=1:5, v=seq(0.1, 0.6, 0.1), col="grey90"),
     cex=as.numeric(GOSharedpathways$Gene.Number) / 4,
     col=colorRamp[cut(as.numeric(GOSharedpathways$p.Value), breaks=50)],
     pch=20, ylab=" ", xlab="Rich Factor", main="Pathways Enriched",
     xlim=c(0, 0.04))
axis(side=2, at=1:5, labels=as.character(GOSharedpathways$GeneSet), las=2, cex.axis=0.5)
axis(side=1, tick=GOSharedpathways[["Rich.Factor"]])
```

### KEGG LA

```{r, eval=FALSE}
LApathways <- read.table("csv/Cytoscape Pathways/LA v LPS.txt", header=TRUE, sep='\t')
LApathways <- LApathways[order(LApathways$p.Value, decreasing=TRUE), ]
## plot x v. y, change size, assign color to p-value
colorRamp <- colorRampPalette(c("turquoise1", "mediumblue"))(50)
par(mar=c(5, 18, 3, 8))
plot(LApathways$Rich.Factor, c(1:28), axes=FALSE,
     panel.first=abline(h=1:28, v=seq(0.1, 0.4, 0.1), col="grey90"),
     cex=as.numeric(LApathways$Gene.Number) / 10,
     col=colorRamp[cut(as.numeric(LApathways$p.Value), breaks=50)],
     pch=20, ylab=" ", xlab="Rich Factor", main="Pathways Enriched",
     xlim=c(0, 0.4))
axis(side=2, at=1:28, labels=as.character(LApathways$Pathway), las=2, cex.axis=0.8)
axis(side=1, tick=LApathways$Rich.Factor, pos=0.4)
```

### GO LA

```{r, eval=FALSE}
GOLApathways <- read.table("csv/Cytoscape Pathways/GO BP LA whole network.txt",
                           header=TRUE, sep='\t')
GOLApathways <- GOLApathways[order(GOLApathways$p.Value, decreasing=TRUE), ]
##plot x v. y, change size, assign color to p-value
colorRamp <- colorRampPalette(c("turquoise1", "mediumblue"))(50)
par(mar=c(5, 20, 3, 8))
plot(GOLApathways$Rich.Factor, c(1:57), axes=F,
     panel.first=abline(h=1:57, v=seq(0.1, 0.5, 0.1), col="grey90"),
     cex=as.numeric(GOLApathways$Gene.Number) / 10,
     col=colorRamp[cut(as.numeric(GOLApathways$p.Value), breaks=50)],
     pch=20, ylab=" ", xlab="Rich Factor", main="Pathways Enriched",
     xlim=c(0, 0.5))
axis(side=2,at=1:57, labels=as.character(GOLApathways$Pathway), las=2, cex.axis=0.8)
axis(side=1, tick=GOLApathways$Rich.Factor, pos=0.4)
```

### GO LP

```{r, eval=FALSE}
GOLPpathways <- read.table("csv/Cytoscape Pathways/GO_BP_LPvLPS.txt", header=TRUE, sep="\t")
GOLPpathways <- GOLPpathways[order(GOLPpathways$p.Value, decreasing=TRUE), ]

colorRamp <- colorRampPalette(c("turquoise1", "mediumblue"))(50)
par(mar=c(5, 18, 3, 8))
plot(GOLPpathways$Rich.Factor, c(1:35), axes=FALSE,
     panel.first=abline(h=1:35, v=seq(0.1, 0.6, 0.1), col="grey90"),
     cex=as.numeric(GOLPpathways$Gene.Number) / 20,
     col=colorRamp[cut(as.numeric(GOLPpathways$p.Value), breaks=50)],
     pch=20, ylab=" ", xlab="Rich Factor", main="Pathways Enriched",
     xlim=c(0, 0.6))
axis(side=2, at=1:35, labels=as.character(GOLPpathways$Pathway), las=2, cex.axis=0.5)
axis(side=1, tick=GOLPpathways$Rich.Factor, pos=0.4)
```

# Volcano plot M_LPS_4 v M_LP_4

## Figure 3B top-left

```{r first_volcano}
res <- read.csv("csv/topTab_M_LPSvM_LP_CDS_limmabatchcorrection_20171120rev.csv",
                header=TRUE)
head(res)
with(res, plot(logFC, -log10(P.Value), cex=0.8,
               pch=20, main="Volcano plot M_LPS v M_LP", xlim=c(-7, 7)))
## Add colored points: red if padj<0.05, orange of log2FC>1, green if both)
with(subset(res, adj.P.Val < 0.05 ), points(logFC, -log10(P.Value), cex=0.8,
                                            pch=20, col="violetred2"))
with(subset(res, abs(logFC) > 1), points(logFC, -log10(P.Value), cex=0.8,
                                         pch=20, col="orange"))
with(subset(res, adj.P.Val < 0.05 & abs(logFC) > 1),
     points(logFC, -log10(P.Value),
            cex=0.8, pch=20, col="turquoise3"))
dev.copy(png, "images/VolcanoPlot_M_LPS_vs_M_LP.png", width=700, height=700)
dev.off()
```

### The number of up significant genes

```{r print_numbers_up1}
sig_up_idx <- res[, "adj.P.Val"] < 0.05 & res[, "logFC"] > 1
sum(sig_up_idx)
sig_up_ids <- res[sig_up_idx, "ID"]
head(sig_up_ids)
m_lps_m_lp_up_ids <- sig_up_ids
```

### The number of down significant genes

```{r print_numbers_down1}
sig_down_idx <- res[, "adj.P.Val"] < 0.05 & res[, "logFC"] < -1
sum(sig_down_idx)
sig_down_ids <- res[sig_down_idx, "ID"]
head(sig_down_ids)
m_lps_m_lp_down_ids <- sig_down_ids
```

# Volcano plot M_LPS_4 v M_LA_4

## Figure 3B top-right

```{r second_volcano}
res <- read.csv(lps_la_table)
head(res)
with(res, plot(logFC, -log10(P.Value), cex=0.8, pch=20,
               main="Volcano plot M_LPS v M_LA", xlim=c(-7, 7)))
## Add colored points: red if padj<0.05, orange of log2FC>1, green if both)
with(subset(res, adj.P.Val < 0.05 ), points(logFC, -log10(P.Value), cex=0.8,
                                            pch=20, col="violetred2"))
with(subset(res, abs(logFC) > 1), points(logFC, -log10(P.Value), cex=0.8,
                                         pch=20, col="orange"))
with(subset(res, adj.P.Val<.05 & abs(logFC) > 1),
     points(logFC, -log10(P.Value),
            cex=0.8, pch=20, col="turquoise3"))
dev.copy(png, "images/VolcanoPlot_M_LPS_v_M_LA.png", width=700, height=700)
dev.off()
```

### The number of up significant genes

```{r print_numbers_up2}
sig_up_idx <- res[, "adj.P.Val"] < 0.05 & res[, "logFC"] > 1
sum(sig_up_idx)
sig_up_ids <- res[sig_up_idx, "ID"]
head(sig_up_ids)
m_lps_m_la_sig_ids <- sig_up_ids
```

### The number of down significant genes

```{r print_numbers_down2}
sig_down_idx <- res[, "adj.P.Val"] < 0.05 & res[, "logFC"] < -1
sum(sig_down_idx)
sig_down_ids <- res[sig_down_idx, "ID"]
head(sig_down_ids)
```

# Volcano plot GM_LPS_4 v GM_LA_4

## Figure 3B bottom-right

```{r third_volcano}
res <- read.csv(gm_lps_la_table)
head(res)
with(res, plot(logFC, -log10(P.Value), cex=0.8, pch=20,
               main="Volcano plot GM_LPS_4 v GM_LA_4", xlim=c(-7, 7)))
## Add colored points: red if padj<0.05, orange of log2FC>1, green if both)
with(subset(res, adj.P.Val < 0.05), points(logFC, -log10(P.Value), cex=0.8,
                                           pch=20, col="violetred2"))
with(subset(res, abs(logFC) > 1), points(logFC, -log10(P.Value), cex=0.8,
                                         pch=20, col="orange"))
with(subset(res, adj.P.Val < 0.05 & abs(logFC) > 1),
     points(logFC, -log10(P.Value), cex=0.8, pch=20, col="turquoise3"))
dev.copy(png, "images/VolcanoPlot GM_LPS v GM_LA", width=700, height=700)
dev.off()
```

### The number of up significant genes

```{r print_numbers_up3}
sig_up_idx <- res[, "adj.P.Val"] < 0.05 & res[, "logFC"] > 1
sum(sig_up_idx)
sig_up_ids <- res[sig_up_idx, "ID"]
head(sig_up_ids)
```

### The number of down significant genes

```{r print_numbers_down3}
sig_down_idx <- res[, "adj.P.Val"] < 0.05 & res[, "logFC"] < -1
sum(sig_down_idx)
sig_down_ids <- res[sig_down_idx, "ID"]
head(sig_down_ids)
```

# Volcano plot GM_LPS_4 v GM_LP_4

## Figure 3B bottom-left

So, it appears I get the inverse of what is in the figure.  Did the contrast get reversed?

I went back to the section which defines this, it has the heading
'GM_LPS_v_GM_LP' and the contrast appears to me to be correct.  At least it is
written like this:

<code>
GM_LPS.GM_LP.contr.mat <- makeContrasts(GM_LPSvGM_LP=((condGM_LPS-condGM_NS)-(condGM_LP-condGM_NS)),
levels=v$design)
</code>

My inclination is therefore that this got flipped?

```{r fourth_volcano}
res <- read.csv(gm_lps_lp_table)
head(res)
with(res, plot(logFC, -log10(P.Value), cex=0.8, pch=20,
               main="Volcano plot GM_LPS v GM_LP", xlim=c(-7, 7)))
## Add colored points: red if padj<0.05, orange of log2FC>1, green if both)
with(subset(res, adj.P.Val < 0.05 ), points(logFC, -log10(P.Value), cex=0.8,
                                            pch=20, col="violetred2"))
with(subset(res, abs(logFC) > 1), points(logFC, -log10(P.Value), cex=0.8,
                                         pch=20, col="orange"))
with(subset(res, adj.P.Val < 0.05 & abs(logFC) > 1),
     points(logFC, -log10(P.Value), cex=0.8, pch=20, col="turquoise3"))
dev.copy(png, "images/VolcanoPlot GM_LPS v GM_LP", width=700, height=700)
dev.off()
```

### The number of up significant genes

```{r print_numbers_up4}
sig_up_idx <- res[, "adj.P.Val"] < 0.05 & res[, "logFC"] > 1
sum(sig_up_idx)
sig_up_ids <- res[sig_up_idx, "ID"]
head(sig_up_ids)
```

### The number of down significant genes

```{r print_numbers_down4}
sig_down_idx <- res[, "adj.P.Val"] < 0.05 & res[, "logFC"] < -1
sum(sig_down_idx)
sig_down_ids <- res[sig_down_idx, "ID"]
head(sig_down_ids)
```

### Figure 2B

I think the code I used for the venn diagrams is in the Rmarkdown file towards
the end.  However, I have 2 files saved for the venn diagrams that are different
from what I have in figure 2B.  I think I need to swap the current 2B with the
venn diagrams I have attached below because the numbers here match with the
volcano plots.  Are these the numbers you are getting also?

#### Attempt 1

This is just pulling the M LP/LPS LA/LPS significant genes.
This appears to not be the answer.

```{r fig2b}
lps_pge_sig <- M_LPS.M_LP.sigGenes
up_idx <- lps_pge_sig[["logFC"]] > 0
down_idx <- lps_pge_sig[["logFC"]] < 0
lps_pge_up_ids <- lps_pge_sig[up_idx, "ID"]
lps_pge_down_ids <- lps_pge_sig[down_idx, "ID"]
lps_a_sig <- M_LPS.M_LA.sigGenes
up_idx <- lps_a_sig[["logFC"]] > 0
down_idx <- lps_a_sig[["logFC"]] < 0
lps_a_up_ids <- lps_a_sig[up_idx, "ID"]
lps_a_down_ids <- lps_a_sig[down_idx, "ID"]

v_data <- list(lps_pge_up_ids, lps_a_up_ids)
v_data <- Venn(v_data, SetNames=c("LPS+PGE2", "LPS+Ado"), numberOfSets=2)
Vennerable::plot(v_data, doWeights=FALSE)

v_data <- list(lps_pge_down_ids, lps_a_down_ids)
v_data <- Venn(v_data, SetNames=c("LPS+PGE2", "LPS+Ado"), numberOfSets=2)
Vennerable::plot(v_data, doWeights=FALSE)
```

#### Attempt 2

Lets assume there is a hint in the text and that I should pull from the volcano
plot numbers for M LP/LPS and M LA/LPS

```{r attempt2}
v_data <- Venn(list(m_lps_m_lp_up_ids, m_lps_m_la_up_ids),
               SetNames=c("LPS+PGE2", "LPS+Ado"))
Vennerable::plot(v_data)

v_data <- Venn(list(m_lps_m_lp_down_ids, m_lps_m_la_down_ids),
               SetNames=c("LPS+PGE2", "LPS+Ado"))
Vennerable::plot(v_data)
```

#### Correlation of logFCs for LP/LPS and LA/LPS

One request in our conversation was to have a calculation of the correlation
coefficient between the LP/LPS and LA/LPS comparisons.

```{r lpla_cor}
m_cor_table <- merge(M_LPS.M_LP.topTab, M_LPS.M_LA.topTab, by="ID")
cor.test(x=m_cor_table[["logFC.x"]], y=m_cor_table[["logFC.y"]])

gm_cor_table <- merge(GM_LPS.GM_LP.topTab, GM_LPS.GM_LA.topTab, by="ID")
cor.test(x=gm_cor_table[["logFC.x"]], y=gm_cor_table[["logFC.y"]])

m_scatter <- plot_linear_scatter(m_cor_table[, c("logFC.x", "logFC.y")])
m_scatter$scatter

gm_scatter <- plot_linear_scatter(gm_cor_table[, c("logFC.x", "logFC.y")])
gm_scatter$scatter
```

```{r saveme}
pander::pander(sessionInfo())
message(paste0("This is hpgltools commit: ", get_git_commit()))

##this_save <- paste0(gsub(pattern="\\.Rmd", replace="", x=rmd_file), "-v", ver, ".rda.xz")
##message(paste0("Saving to ", this_save))
##tmp <- sm(saveme(filename=this_save))
```
