---
title: "Experimenting with some scRNASeq data."
author: "atb abelew@gmail.com"
date: "`r Sys.Date()`"
output:
  html_document:
    code_download: true
    code_folding: show
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: tango
    keep_md: false
    mode: selfcontained
    number_sections: true
    self_contained: true
    theme: readable
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
  rmdformats::readthedown:
    code_download: true
    code_folding: show
    df_print: paged
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: tango
    width: 300
    keep_md: false
    mode: selfcontained
    toc_float: true
  BiocStyle::html_document:
    code_download: true
    code_folding: show
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: tango
    keep_md: false
    mode: selfcontained
    toc_float: true
---

<style type="text/css">
body, td {
  font-size: 16px;
}
code.r{
  font-size: 16px;
}
pre {
 font-size: 16px
}
</style>

```{r options, include=FALSE}
library("hpgltools")
tt <- devtools::load_all("/data/hpgltools")
knitr::opts_knit$set(width=120,
                     progress=TRUE,
                     verbose=TRUE,
                     echo=TRUE)
knitr::opts_chunk$set(error=TRUE,
                      dpi=96)
old_options <- options(digits=4,
                       stringsAsFactors=FALSE,
                       knitr.duplicate.label="allow")
ggplot2::theme_set(ggplot2::theme_bw(base_size=10))
rundate <- format(Sys.Date(), format="%Y%m%d")
previous_file <- "index.Rmd"
ver <- "202004"

##tmp <- sm(loadme(filename=paste0(gsub(pattern="\\.Rmd", replace="", x=previous_file), "-v", ver, ".rda.xz")))
rmd_file <- "scrnaseq_analyses.Rmd"
```

# Testing seurat

```{r seurat_loading}
library(Seurat)
library(ggplot2)
## library(SeuratData)
## The Seurat vignettes show how to load hdf5 files and the metadata required.
## They are using a pacreas dataset, I presume it is the 'panc8' data in SeuratData.
## Thus I will poke at that to see if I can learn about how they handle the metadata.

la_data <- Seurat::Read10X("khamidza_158974/KH_LA/outs/filtered_feature_bc_matrix")
lp_data <- Seurat::Read10X("khamidza_158974/KH_LP/outs/filtered_feature_bc_matrix")
lps_data <- Seurat::Read10X("khamidza_158974/KH_LPS/outs/filtered_feature_bc_matrix")
ns_data <- Seurat::Read10X("khamidza_158974/KH_NS/outs/filtered_feature_bc_matrix")

la <- Seurat::CreateSeuratObject(la_data, project="la")
lp <- Seurat::CreateSeuratObject(lp_data, project="lp")
lps <- Seurat::CreateSeuratObject(lps_data, project="lps")
ns <- Seurat::CreateSeuratObject(ns_data, project="ns")

all <- merge(la, lp)
all <- merge(all, lps)
all <- merge(all, ns)

## I think this is redundant, but interesting for me to understand sc metadata
cluster_letters <- as.factor(LETTERS[Idents(object=all)])
names(cluster_letters) <- colnames(x=all)

stimulated <- as.character(cluster_letters)
ns_idx <- stimulated == "D"
stimulated[ns_idx] <- "unstimulated"
stimulated[!ns_idx] <- "stimulated"
stimulated <- as.factor(stimulated)

stimulation <- as.character(cluster_letters)
ns_idx <- stimulation == "D"
stimulation[ns_idx] <- "Unstimulated"
la_idx <- stimulation == "A"
stimulation[la_idx] <- "L+Ado"
lp_idx <- stimulation == "B"
stimulation[lp_idx] <- "L+PGE2"
lps_idx <- stimulation == "C"
stimulation[lps_idx] <- "LPS"
stimulation <- as.factor(stimulation)

all <- AddMetaData(
  object=all,
  metadata=cluster_letters,
  col.name="cluster_letters")
all <- AddMetaData(
    object=all,
    metadata=stimulated,
    col.name="stimulatedp")
all <- AddMetaData(
    object=all,
    metadata=stimulation,
    col.name="stimulation")

all[["percent_mt"]] <- PercentageFeatureSet(all, pattern="^MT-")
VlnPlot(all, features="nFeature_RNA", pt.size=0.1)
VlnPlot(all, features="percent_mt", pt.size=0.1)
VlnPlot(all, features="nCount_RNA", pt.size=0.1)

all <- NormalizeData(object=all)
all <- FindVariableFeatures(object=all)
all <- ScaleData(object=all)
all <- RunPCA(object=all)
all <- FindNeighbors(object=all)
all <- FindClusters(object=all)
all <- RunTSNE(object=all)
all <- RunUMAP(all, reduction="pca", dims=1:20)
DimPlot(object=all, reduction="tsne")
plotted <- DimPlot(all, reduction="umap", group.by="cluster_letters", label=TRUE)
plotted

## Try to set some metrics to drop crappy stuff.
FeatureScatter(all, feature1="nCount_RNA", feature2="percent_mt")
## Suggests that we want ~  < 15% mt
## Suggests we want > 1000 nCounts
FeatureScatter(all, feature1="nCount_RNA", feature2="nFeature_RNA")
## Suggests we want > 200 nFeature
FeatureScatter(all, feature1="percent_mt", feature2="nFeature_RNA")
all_sub <- subset(all,
                  subset=nFeature_RNA > 200 & percent_mt < 15)
all_sub <- NormalizeData(object=all_sub)
all_sub <- FindVariableFeatures(object=all_sub)
most_var <- head(VariableFeatures(all_sub), 30)
variable_plot <- VariableFeaturePlot(all_sub)
variable_plot <- LabelPoints(plot=variable_plot, points=most_var, repel=TRUE)
variable_plot
all_sub <- ScaleData(object=all_sub)
all_sub <- RunPCA(object=all_sub)
VizDimLoadings(all_sub, dims=1:2, reduction="pca")
all_sub <- JackStraw(all_sub, num.replicate=100)
all_sub <- ScoreJackStraw(all_sub, dims=1:20)
JackStrawPlot(all_sub, dims=1:15)
ElbowPlot(all_sub)
all_sub <- FindNeighbors(object=all_sub)
all_sub <- FindClusters(object=all_sub)
all_sub <- RunTSNE(object=all_sub)
all_sub <- RunUMAP(all_sub, reduction="pca", dims=1:20)
DimPlot(object=all_sub, reduction="tsne")
plotted <- DimPlot(all_sub, reduction="umap", group.by="stimulation", label=TRUE)
plotted
plotted <- DimPlot(all_sub, reduction="pca", label=TRUE)
plotted
```

```{r markers}
DefaultAssay(all) <- "RNA"
markers <- FindConservedMarkers(all_sub, ident.1=4, grouping.var="stimulation", verbose=TRUE)

## Note that I renamed them according to Dave's suggestion.
lps_vs_unstim <- FindMarkers(all_sub,
                             ident.1="LPS",
                             ident.2="Unstimulated",
                             group.by="stimulation",
                             min.pct=0.25)
head(lps_vs_unstim, n=20)
ade_vs_unstim <- FindMarkers(all_sub,
                             ident.1="L+Ado",
                             ident.2="Unstimulated",
                             group.by="stimulation",
                             min.pct=0.25)
head(ade_vs_unstim, n=20)
pge_vs_unstim <- FindMarkers(all_sub,
                             ident.1="L+PGE2",
                             ident.2="Unstimulated",
                             group.by="stimulation",
                             min.pct=0.25)
head(pge_vs_unstim, n=20)

ado_lps <- FindMarkers(all_sub,
                       ident.1="L+Ado",
                       ident.2="LPS",
                       group.by="stimulation", min.pct=0.25)
head(ado_lps, n=20)

pge2_lps <- FindMarkers(all_sub,
                        ident.1="L+PGE2",
                        ident.2="LPS",
                        group.by="stimulation", min.pct=0.25)
head(pge2_lps, n=20)
```

```{r features}
FeaturePlot(all_sub, features=c("THBS1"),
            split.by="stimulation", max.cutoff=3, cols=c("darkgreen", "darkred"))
FeaturePlot(all_sub, features=c("PLEK"),
            split.by="stimulation", max.cutoff=3, cols=c("darkgreen", "darkred"))
FeaturePlot(all_sub, features=c("FNIP2"),
            split.by="stimulation", max.cutoff=3, cols=c("darkgreen", "darkred"))
FeaturePlot(all_sub, features=c("G0S2"),
            split.by="stimulation", max.cutoff=3, cols=c("darkgreen", "darkred"))
```

## Figure 2A

This looks to me like a bar plot of the top 10 up/down genes.  But I have no
clue where these numbers are coming from, a range of -5 < x < 5 logFC just does
not seem to me to exist in this data.  The range of logFCs on my sheet goes from
-1.5 < x < 1.7.  In addition, I do not see where in the Seurat data structures
the error bars are coming from.

```{r fig2a}
ado_lps <- FindMarkers(all_sub, test.use="DESeq2",
                       ident.1="L+Ado",
                       ident.2="LPS",
                       group.by="stimulation", min.pct=0.25)

```

## Figure 2B

Reading from the text, this appears to me to be from the total cell RNASeq?  No,
that cannot be true, the bulk data did not compare these things.  It must be
this data.

I did generate a couple of Venns using the scRNA data and got similar numbers in
the opposite orientation.

```{r venn}
ado <- rownames(ado_lps)
ado_up <- ado[ado_lps[["avg_logFC"]] > 0]
ado_down <- ado[ado_lps[["avg_logFC"]] < 0]
pge2 <- rownames(pge2_lps)
pge_up <- pge2[pge2_lps[["avg_logFC"]] > 0]
pge_down <- pge2[pge2_lps[["avg_logFC"]] < 0]

library(Vennerable)
ups <- list(ado_up, pge_up)
up_data <- Venn(ups, SetNames=c("LPS+Ado", "LPS+PGE2"), numberOfSets=2)
plot(up_data, doWeights=FALSE)

downs <- list(ado_down, pge_down)
v_data <- Venn(downs, SetNames=c("LPS+Ado", "LPS+PGE2"), numberOfSets=2)
plot(v_data, doWeights=FALSE)
```

## Figure 2C

```{r fig2c}
plotted <- DimPlot(all_sub, reduction="umap", group.by="stimulation", label=TRUE)
plotted
```

## Figure 5A

```{r fig5a}
thbs1 <- VlnPlot(all, features="THBS1", group.by="stimulation", pt.size=0.1)
thbs1
box <- ggplot(data=thbs1$data, aes(x=ident, y=THBS1)) +
  geom_boxplot(notch=TRUE)
box
input <- thbs1$data
thbs1_ggstats <- ggstatsplot::ggbetweenstats(
                                  data=input, x=ident, y=THBS1,
                                  notch=TRUE, mean.ci=TRUE, k=3,
                                  pairwise.comparisons=TRUE)
thbs1_ggstats
##  pairwise.comparisons=TRUE, # display significant pairwise comparisons
##  p.adjust.method="bonferroni", # method for adjusting p-values for multiple comparisons
##  ## adding new components to `ggstatsplot` default
##  ##ggplot.component = list(ggplot2::scale_y_continuous(sec.axis = ggplot2::dup_axis())),
##  k=3, title.prefix="THBS1", palette="default_jama",
##  package="ggsci", messages=TRUE, plotgrid.args=list(nrow=2))

vegfa <- VlnPlot(all, features="VEGFA", group.by="stimulation", pt.size=0.1)
vegfa
vegfa_ggstats <- ggstatsplot::ggbetweenstats(
                                  data=vegfa$data, x=ident, y=VEGFA,
                                  notch=TRUE, mean.ci=TRUE, k=3,
                                  pairwise.comparisons=TRUE)
vegfa_ggstats

cd300e <- VlnPlot(all, features="CD300E", group.by="stimulation", pt.size=0.1)
cd300e_ggstats <- ggstatsplot::ggbetweenstats(
                                   data=cd300e$data, x=ident, y=CD300E,
                                   notch=TRUE, mean.ci=TRUE, k=3,
                                   pairwise.comparisons=TRUE)
cd300e_ggstats

plaur <- VlnPlot(all, features="PLAUR", group.by="stimulation", pt.size=0.1)
plaur_ggstats <- ggstatsplot::ggbetweenstats(
                                   data=plaur$data, x=ident, y=PLAUR,
                                   notch=TRUE, mean.ci=TRUE, k=3,
                                   pairwise.comparisons=TRUE)
plaur_ggstats
```

```{r saveme}
pander::pander(sessionInfo())
message(paste0("This is hpgltools commit: ", get_git_commit()))
this_save <- paste0(gsub(pattern="\\.Rmd", replace="", x=rmd_file), "-v", ver, ".rda.xz")
message(paste0("Saving to ", this_save))
tmp <- sm(saveme(filename=this_save))
```

```{r loadme, eval=FALSE}
this_save <- paste0(gsub(pattern="\\.Rmd", replace="", x=rmd_file), "-v", ver, ".rda.xz")
loaded <- loadme(filename=this_save)
```
