---
title: "Human Macrophages: atb version of M-CSF v GM-CSF: LPS, LPS+Adenosine, LPS+PGE2"
author: "Kajal Hamidzadeh and atb"
date: "`r Sys.Date()`"
output:
  html_document:
    code_download: true
    code_folding: show
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: tango
    keep_md: false
    mode: selfcontained
    number_sections: true
    self_contained: true
    theme: readable
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
  rmdformats::readthedown:
    code_download: true
    code_folding: show
    df_print: paged
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: tango
    width: 300
    keep_md: false
    mode: selfcontained
    toc_float: true
  BiocStyle::html_document:
    code_download: true
    code_folding: show
    fig_caption: true
    fig_height: 7
    fig_width: 7
    highlight: tango
    keep_md: false
    mode: selfcontained
    toc_float: true
---

<style type="text/css">
body, td {
  font-size: 16px;
}
code.r{
  font-size: 16px;
}
pre {
 font-size: 16px
}
</style>

```{r options, include=FALSE}
library("hpgltools")
tt <- devtools::load_all("/data/hpgltools")
knitr::opts_knit$set(width=120,
                     progress=TRUE,
                     verbose=TRUE,
                     echo=TRUE)
knitr::opts_chunk$set(error=TRUE,
                      dpi=96)
old_options <- options(digits=4,
                       stringsAsFactors=FALSE,
                       knitr.duplicate.label="allow")
ggplot2::theme_set(ggplot2::theme_bw(base_size=10))
rundate <- format(Sys.Date(), format="%Y%m%d")
previous_file <- ""
ver <- "20200330"

##tmp <- sm(loadme(filename=paste0(gsub(pattern="\\.Rmd", replace="", x=previous_file), "-v", ver, ".rda.xz")))
##rmd_file <- "03_expression_infection_20180822.Rmd"
```

In this version of the worksheet, I am hoping to perform basically the same
analyses, but do it in a fashion which is more in my own style.

# Annotation data

Collect the human annotation data using biomaRt.

```{r annotation}
gene_info <- load_biomart_annotations(host="useast.ensembl.org")$annotation
rownames(gene_info) <- make.names(gene_info[["ensembl_gene_id"]], unique=TRUE)
tx_gene_map <- gene_info[, c("ensembl_transcript_id", "ensembl_gene_id")]
```

# Experimental design

I am going to use Kajal's sample sheet without modification.

```{r expt_design}
design <- read.table("sample_sheets/MetaData only 4 hour.txt", header=TRUE, sep='\t')
design[["Patient"]] <- as.factor(design[["Patient"]])
design[["Stimulation"]] <- as.factor(design[["Stimulation"]])
design[["Batch"]] <- as.factor(design[["Batch"]])
design[["Growth"]] <- as.factor(design[["Growth"]])
files <- file.path("kallisto abundance files/", design$HPGL.Identifier, "abundance.tsv")
names(files) <- paste0("HPGL09", c(12:31, 42:60))
rownames(design) <- design[[1]]
design[["condition"]] <- design[["Stimulation"]]
design[["file"]] <- glue::glue("preprocessing/{rownames(design)}/abundance.tsv")
colnames(design) <- tolower(colnames(design))
design[["gp"]] <- as.factor(glue("{design[['growth']]}_{design[['stimulation']]}"))

## Set up a column called stim_pred which is a predicate of stimulated vs. unstimulated samples.
design[["stim_pred"]] <- "stimulated"
ns_idx <- design[["stimulation"]] == "NS"
design[ns_idx, "stim_pred"] <- "unstimulated"
```

# Create expressionset

We have some annotation data and experimental metadata.

```{r create_expt}
hs_expt <- create_expt(metadata=design, gene_info=gene_info, tx_gene_map=tx_gene_map)
hs_expt <- set_expt_batches(hs_expt, fact="growth")
stim_expt <- subset_expt(hs_expt, subset="stimulation!='NS'")
```

# Write out the expressionset data

We can write out the data to an excel file in the hopes that it will prove useful.

```{r write_expt, fig.show="hide"}
written <- write_expt(hs_expt, batch="raw",
                      excel=glue::glue("excel/hs_expt-v{ver}.xlsx"))
```

There is an important caveat, this is not taking into account the patient effects.

## Show some written plots

```{r write_plots}
written$legend_plot
written$raw_libsize
written$raw_density
written$raw_boxplot
written$norm_nonzero
written$norm_corheat
written$norm_disheat
written$norm_pca
```

# Consider different models for the data

We may wish to lower the variance from the patients and/or the GM/M effects.

## Current state with sva

```{r current_sva}
hs_batch <- normalize_expt(hs_expt, transform="log2", convert="cpm",
                           norm="quant", filter=TRUE, batch="svaseq")
plot_pca(hs_batch)$plot
```

## Current state with residual-based batch adjustment

Because we are explicitly removing the effect of GM/M, the patient effect really
becomes apparent.

```{r current_limma}
hs_batch <- normalize_expt(hs_expt, transform="log2", convert="cpm",
                           norm="quant", filter=TRUE, batch="limmaresid")
plot_pca(hs_batch)$plot
```

## Set batch to patient and repeat sva

The picture with sva should be the same as the first plot, just with 6 shapes
instead of two.

```{r patient_sva}
hs_pat <- set_expt_batches(hs_expt, fact="patient")
hs_batch <- normalize_expt(hs_pat, transform="log2", convert="cpm",
                           norm="quant", filter=TRUE, batch="svaseq")
plot_pca(hs_batch)$plot
```

## Patient batch and limma

This picture should be different, and should show us the M/GM effect as opposed
to the patient effect.

```{r patient_limma}
hs_batch <- normalize_expt(hs_pat, transform="log2", convert="cpm",
                           norm="quant", filter=TRUE, batch="limmaresid")
plot_pca(hs_batch)$plot
```

## Combine growth and stimulation and repeat sva

I am not sure what this will look like.  Since two aspects of the data are in
the condition portion of the model matrix, I think it should look different.
When I created the design matrix, I made a column for this purpose; I called it
'gp', but honestly I don't remember why...

```{r gs_sva}
hs_gs <- set_expt_conditions(hs_pat, fact="gp")
hs_batch <- normalize_expt(hs_gs, transform="log2", convert="cpm",
                           norm="quant", filter=TRUE, batch="svaseq")
plot_pca(hs_batch)$plot
```

It seems to me that is primarily showing us differences between M/GM.

What about limma?

```{r gs_limma}
hs_batch <- normalize_expt(hs_gs, transform="log2", convert="cpm",
                           norm="quant", filter=TRUE, batch="limma")
plot_pca(hs_batch)$plot
```

Same deal, just more.  This might be the moment to reconsider the fact that
Kajal's work focused only on the M/GM samples and AFAICT ignored the
non-stimulated samples.

```{r test}
tmp <- subset_expt(hs_gs, subset='stimulation!="NS"')
hs_batch <- normalize_expt(tmp, transform="log2", convert="cpm",
                           norm="quant", filter=TRUE, batch="sva")
plot_pca(hs_batch)$plot

tmp <- subset_expt(hs_gs, subset='stimulation!="NS"')
hs_ruv <- normalize_expt(tmp, transform="log2", convert="cpm",
                         norm="quant", filter=TRUE, batch="ruv_empirical")
plot_pca(hs_ruv)$plot

hs_limma <- normalize_expt(tmp, transform="log2", convert="cpm",
                           norm="quant", filter=TRUE, batch="limmaresid")
plot_pca(hs_limma)$plot
```

### My Figure 3A, sva

Here are a few versions of what figure 3a might look like in my world.

```{r fig3asva1}
plot_pca(hs_batch, plot_labels=FALSE)$plot
plot_pca(hs_ruv, plot_labels=FALSE)$plot
plot_pca(hs_limma, plot_labels=FALSE)$plot
```

Interesting, I did not put them all into the test block above, but I tried out a
bunch of small changes to the model and adjusters.  I think I learned one
primary lesson: patient number 3 is a bit weird.  I think I might suggest
removing this person from the data.

The next lesson I learned is that LPS is way different than LA/LP, something
which I kind of knew from other work, but worth remembering.

# Differential expression

There are a few ways to consider differential expression for this data.  In all
cases I think it is safe to assume that we wish to use patient as the batch
factor/surrogate variable.

With that in mind, here are the factors of the data to which we have usable
variance/experimental design:

1.  Stimulated vs. Unstimulated:  This I think has the most variance in the
    data, even including patient.  We can access this by putting stimulation
    state (yes/no) in the model and just running sva against everything else, or by
    having a model like  "~ stimulation_binary + patient + growth"
2.  Stimulation types:  If we want to consider all LPS vs. all LP etc... we can
    do that in a similar fashion, by either putting stimulation state
    (LPS/LP/etc) in the model and just running sva.  Conversely we could do
    "~ stimulation_state + patient + growth" in the model.
3.  Growth condition:  Ibid, except "~ growth + patient + stimulation"
4.  Growth+Stimulation:  This is the focus of Kajal's worksheet I think and may
    be repeated with "~ gp + patient"

Before I run these, lets look at the variance in the data and make sure I am not
full of crap.

```{r varpart}
hs_vpin <- normalize_expt(hs_expt, convert="cpm", filter=TRUE)
hs_varpart <- simple_varpart(hs_vpin, factors=c("stimulation", "growth", "patient"),
                             chosen_factor="patient", do_fit=TRUE)
hs_varpart$partition_plot
top_40_stimulation <- hs_varpart$percent_plot
## Now show a variance boxplot for the chosen batch factor (patient)
hs_varpart$stratify_batch_plot
hs_varpart$stratify_condition_plot

percent_growth <- replot_varpart_percent(hs_varpart, column="growth")
percent_growth$plot

percent_patient <- replot_varpart_percent(hs_varpart, column="patient")
percent_patient$plot

percent_unknown <- replot_varpart_percent(hs_varpart, column="Residuals")
percent_unknown$plot
```
## Perform some DE

### Patient as batch, growth and stimulation as condition, sva

```{r de_patbatch_gscond_batch, fig.show="hide"}
hs_filt <- normalize_expt(hs_expt, filter=TRUE)
pat_gs_sva <- set_expt_conditions(hs_filt, fact="gp")
pat_gs_sva <- set_expt_batches(pat_gs_sva, fact="patient")
pat_gs_sva_de <- all_pairwise(pat_gs_sva, model_batch="sva")

keepers <- list(
    ## GM against unstimulated
    "GM_LPS_vs_GM_NS" = c("GM_LPS", "GM_NS"),
    "GM_LP_vs_GM_NS" = c("GM_LP", "GM_NS"),
    "GM_LA_vs_GM_NS" = c("GM_LA", "GM_NS"),
    ## M against unstimulated
    "M_LPS_vs_M_NS" = c("M_LPS", "M_NS"),
    "M_LP_vs_M_NS" = c("M_LP", "M_NS"),
    "M_LA_vs_M_NS" = c("M_LA", "M_NS"),
    ## GM against LPS
    "GM_LP_vs_GM_LPS" = c("GM_LP", "GM_LPS"),
    "GM_LA_vs_GM_LPS" = c("GM_LA", "GM_LPS"),
    ## M against LPS
    "M_LP_vs_M_LPS" = c("M_LP", "M_LPS"),
    "M_LA_vs_M_LPS" = c("M_LA", "M_LPS"),
    ## GM, LA vs LP
    "GM_LP_vs_GM_LA" = c("GM_LP", "GM_LA"),
    ## M, LA vs LP
    "M_LP_vs_M_LA" = c("M_LP", "M_LA"),
    ## Last, each M vs GM
    "GM_NS_vs_M_NS" = c("GM_NS", "M_NS"),
    "GM_LPS_vs_M_LPS" = c("GM_LPS", "M_LPS"),
    "GM_LP_vs_M_LP" = c("GM_LP", "M_LP"),
    "GM_LA_vs_M_LA" = c("GM_LA", "M_LA"))
pat_gs_sva_tables <- combine_de_tables(
    pat_gs_sva_de, keepers=keepers,
    excel=glue::glue("excel/pat_gs_sva_tables-v{ver}.xlsx"))
pat_gs_sva_sig <- extract_significant_genes(
    pat_gs_sva_tables,
    excel=glue::glue("excel/pat_gs_sva_sig-v{ver}.xlsx"))
```

### Patient as batch, growth and stimulation as condition, batch in model.

```{r de_patbatch_gscond_sva, fig.show="hide"}
pat_gs_batch <- set_expt_conditions(hs_filt, fact="gp")
pat_gs_batch <- set_expt_batches(pat_gs_batch, fact="patient")
pat_gs_batch_de <- all_pairwise(pat_gs_batch, model_batch=TRUE)

pat_gs_batch_tables <- combine_de_tables(
    pat_gs_batch_de, keepers=keepers,
    excel=glue::glue("excel/pat_gs_batch_tables-v{ver}.xlsx"))
pat_gs_batch_sig <- extract_significant_genes(
    pat_gs_batch_tables,
    excel=glue::glue("excel/pat_gs_batch_sig-v{ver}.xlsx"))


```

## Compare sva/batch in model

```{r compare_de}
comp <- compare_de_results(pat_gs_sva_tables, pat_gs_batch_tables)
## Look at the logFC comparisons:
comp$lfc_heat
## Look at the p-value comparisons:
comp$p_heat
## It appears edgeR is a bit more sensitive to changes in the model.
```

# Perform some ontology searches

It appears that I crashed the gProfiler web server by sending in my various
searches.  So I will leave these off for the moment and replace them with some
clusterProfiler searches.

## GM Comparisons

### GM LPS vs GM NS

```{r}
table <- "GM_LPS_vs_GM_NS"
up <- pat_gs_sva_sig[["deseq"]][["ups"]][[table]]
down <- pat_gs_sva_sig[["deseq"]][["downs"]][[table]]
de <- pat_gs_sva_tables[["data"]][[table]]
```

```{r gprofiler1, eval=FALSE}
ont_up <- simple_gprofiler(up)
ont_up[["pvalue_plots"]][["mfp_plot_over"]]
ont_up[["pvalue_plots"]][["bpp_plot_over"]]
ont_up[["pvalue_plots"]][["kegg_plot_over"]]
ont_up[["pvalue_plots"]][["reactome_plot_over"]]
ont_up[["pvalue_plots"]][["hp_plot_over"]]

ont_down <- simple_gprofiler(down)
ont_down[["pvalue_plots"]][["mfp_plot_over"]]
ont_down[["pvalue_plots"]][["bpp_plot_over"]]
ont_down[["pvalue_plots"]][["kegg_plot_over"]]
ont_down[["pvalue_plots"]][["reactome_plot_over"]]
ont_down[["pvalue_plots"]][["hp_plot_over"]]
```

```{r cp1}
ont_up <- simple_clusterprofiler(sig_genes=up, de_table=de,
                                 do_david=TRUE, david_user="abelew@umd.edu",
                                 fc_column="deseq_logfc", orgdb="org.Hs.eg.db")
ont_up$plots$ego_sig_mf
ont_up$plots$ego_sig_bp
ont_up$plots$dot_sig_mf
ont_up$plots$dot_sig_bp
ont_up$plots$map_sig_mf
ont_up$plots$map_sig_bp

ont_down <- simple_clusterprofiler(sig_genes=down, de_table=de,
                                 do_david=TRUE, david_user="abelew@umd.edu",
                                 fc_column="deseq_logfc", orgdb="org.Hs.eg.db")
## No significant hits going down
ont_down$plots$ego_all_mf
ont_down$plots$ego_all_bp
ont_down$plots$dot_all_mf
ont_down$plots$dot_all_bp
ont_down$plots$map_all_mf
ont_down$plots$map_all_bp
```

### GM LP vs GM NS

```{r}
table <- "GM_LP_vs_GM_NS"
up <- pat_gs_sva_sig[["deseq"]][["ups"]][[table]]
down <- pat_gs_sva_sig[["deseq"]][["downs"]][[table]]
de <- pat_gs_sva_tables[["data"]][[table]]
```

```{r gp2, eval=FALSE}
ont_up <- simple_gprofiler(up)
ont_up[["pvalue_plots"]][["mfp_plot_over"]]
ont_up[["pvalue_plots"]][["bpp_plot_over"]]
ont_up[["pvalue_plots"]][["kegg_plot_over"]]
ont_up[["pvalue_plots"]][["reactome_plot_over"]]
ont_up[["pvalue_plots"]][["hp_plot_over"]]

ont_down <- simple_gprofiler(down)
ont_down[["pvalue_plots"]][["mfp_plot_over"]]
ont_down[["pvalue_plots"]][["bpp_plot_over"]]
ont_down[["pvalue_plots"]][["kegg_plot_over"]]
ont_down[["pvalue_plots"]][["reactome_plot_over"]]
ont_down[["pvalue_plots"]][["hp_plot_over"]]
```

```{r cp2}
ont_up <- simple_clusterprofiler(sig_genes=up, de_table=de,
                                 do_david=TRUE, david_user="abelew@umd.edu",
                                 fc_column="deseq_logfc", orgdb="org.Hs.eg.db")
ont_up$plots$ego_sig_mf
ont_up$plots$ego_sig_bp
ont_up$plots$dot_sig_mf
ont_up$plots$dot_sig_bp
ont_up$plots$map_sig_mf
ont_up$plots$map_sig_bp

ont_down <- simple_clusterprofiler(sig_genes=down, de_table=de,
                                 do_david=TRUE, david_user="abelew@umd.edu",
                                 fc_column="deseq_logfc", orgdb="org.Hs.eg.db")
## No significant hits going down
ont_down$plots$ego_all_mf
ont_down$plots$ego_all_bp
ont_down$plots$dot_all_mf
ont_down$plots$dot_all_bp
ont_down$plots$map_all_mf
ont_down$plots$map_all_bp
```

### GM LA vs GM NS

```{r}
table <- "GM_LA_vs_GM_NS"
up <- pat_gs_sva_sig[["deseq"]][["ups"]][[table]]
down <- pat_gs_sva_sig[["deseq"]][["downs"]][[table]]
de <- pat_gs_sva_tables[["data"]][[table]]
```

```{r gp3, eval=FALSE}
ont_up <- simple_gprofiler(up)
ont_up[["pvalue_plots"]][["mfp_plot_over"]]
ont_up[["pvalue_plots"]][["bpp_plot_over"]]
ont_up[["pvalue_plots"]][["kegg_plot_over"]]
ont_up[["pvalue_plots"]][["reactome_plot_over"]]
ont_up[["pvalue_plots"]][["hp_plot_over"]]

ont_down <- simple_gprofiler(down)
ont_down[["pvalue_plots"]][["mfp_plot_over"]]
ont_down[["pvalue_plots"]][["bpp_plot_over"]]
ont_down[["pvalue_plots"]][["kegg_plot_over"]]
ont_down[["pvalue_plots"]][["reactome_plot_over"]]
ont_down[["pvalue_plots"]][["hp_plot_over"]]
```

```{r cp3}
ont_up <- simple_clusterprofiler(sig_genes=up, de_table=de,
                                 do_david=TRUE, david_user="abelew@umd.edu",
                                 fc_column="deseq_logfc", orgdb="org.Hs.eg.db")
ont_up$plots$ego_sig_mf
ont_up$plots$ego_sig_bp
ont_up$plots$dot_sig_mf
ont_up$plots$dot_sig_bp
ont_up$plots$map_sig_mf
ont_up$plots$map_sig_bp

ont_down <- simple_clusterprofiler(sig_genes=down, de_table=de,
                                   do_david=TRUE, david_user="abelew@umd.edu",
                                   fc_column="deseq_logfc", orgdb="org.Hs.eg.db")
## No significant hits going down
ont_down$plots$ego_all_mf
ont_down$plots$ego_all_bp
ont_down$plots$dot_all_mf
ont_down$plots$dot_all_bp
ont_down$plots$map_all_mf
ont_down$plots$map_all_bp
```

## GM Comparisons against LPS

### GM LP vs GM LPS

```{r}
table <- "GM_LP_vs_GM_LPS"
up <- pat_gs_sva_sig[["deseq"]][["ups"]][[table]]
down <- pat_gs_sva_sig[["deseq"]][["downs"]][[table]]
de <- pat_gs_sva_tables[["data"]][[table]]
```


```{r gp4, eval=FALSE}
ont_up <- simple_gprofiler(up)
ont_up[["pvalue_plots"]][["mfp_plot_over"]]
ont_up[["pvalue_plots"]][["bpp_plot_over"]]
ont_up[["pvalue_plots"]][["kegg_plot_over"]]
ont_up[["pvalue_plots"]][["reactome_plot_over"]]
ont_up[["pvalue_plots"]][["hp_plot_over"]]

ont_down <- simple_gprofiler(down)
ont_down[["pvalue_plots"]][["mfp_plot_over"]]
ont_down[["pvalue_plots"]][["bpp_plot_over"]]
ont_down[["pvalue_plots"]][["kegg_plot_over"]]
ont_down[["pvalue_plots"]][["reactome_plot_over"]]
ont_down[["pvalue_plots"]][["hp_plot_over"]]
```

```{r cp4}
ont_up <- simple_clusterprofiler(sig_genes=up, de_table=de,
                                 do_david=TRUE, david_user="abelew@umd.edu",
                                 fc_column="deseq_logfc", orgdb="org.Hs.eg.db")
ont_up$plots$ego_sig_mf
ont_up$plots$ego_sig_bp
ont_up$plots$dot_sig_mf
ont_up$plots$dot_sig_bp
ont_up$plots$map_sig_mf
ont_up$plots$map_sig_bp

ont_down <- simple_clusterprofiler(sig_genes=down, de_table=de,
                                   do_david=TRUE, david_user="abelew@umd.edu",
                                   fc_column="deseq_logfc", orgdb="org.Hs.eg.db")
## No significant hits going down
ont_down$plots$ego_all_mf
ont_down$plots$ego_all_bp
ont_down$plots$dot_all_mf
ont_down$plots$dot_all_bp
ont_down$plots$map_all_mf
ont_down$plots$map_all_bp
```

### GM LA vs GM LPS

```{r}
table <- "GM_LA_vs_GM_LPS"
up <- pat_gs_sva_sig[["deseq"]][["ups"]][[table]]
down <- pat_gs_sva_sig[["deseq"]][["downs"]][[table]]
de <- pat_gs_sva_tables[["data"]][[table]]
```

```{r gp5, eval=FALSE}
ont_up <- simple_gprofiler(up)
## No hits!
```

### GM LA vs GM LP

This does not get any useful results.

```{r, eval=FALSE}
up <- pat_gs_sva_sig[["deseq"]][["ups"]][["GM_LP_vs_GM_LA"]]
down <- pat_gs_sva_sig[["deseq"]][["downs"]][["GM_LP_vs_GM_LA"]]

ont_up <- simple_gprofiler(up)
ont_up[["pvalue_plots"]][["mfp_plot_over"]]
ont_up[["pvalue_plots"]][["bpp_plot_over"]]
ont_up[["pvalue_plots"]][["kegg_plot_over"]]
ont_up[["pvalue_plots"]][["reactome_plot_over"]]
ont_up[["pvalue_plots"]][["hp_plot_over"]]

## Down only had 2 genes.
```

```{r cp5}
ont_up <- simple_clusterprofiler(sig_genes=up, de_table=de,
                                 do_david=TRUE, david_user="abelew@umd.edu",
                                 fc_column="deseq_logfc", orgdb="org.Hs.eg.db")
ont_up$plots$ego_sig_mf
ont_up$plots$ego_sig_bp
ont_up$plots$dot_sig_mf
ont_up$plots$dot_sig_bp
ont_up$plots$map_sig_mf
ont_up$plots$map_sig_bp

ont_down <- simple_clusterprofiler(sig_genes=down, de_table=de,
                                   do_david=TRUE, david_user="abelew@umd.edu",
                                   fc_column="deseq_logfc", orgdb="org.Hs.eg.db")
## No significant hits going down
ont_down$plots$ego_all_mf
ont_down$plots$ego_all_bp
ont_down$plots$dot_all_mf
ont_down$plots$dot_all_bp
ont_down$plots$map_all_mf
ont_down$plots$map_all_bp
```

## M Comparisons

### M LPS vs M NS

This search crashed the gProfiler server, so I will stop it at least for the moment.

```{r}
table <- "M_LPS_vs_M_NS"
up <- pat_gs_sva_sig[["deseq"]][["ups"]][[table]]
down <- pat_gs_sva_sig[["deseq"]][["downs"]][[table]]
de <- pat_gs_sva_tables[["data"]][[table]]
```

```{r gp6, eval=FALSE}
ont_up <- simple_gprofiler(up)
ont_up[["pvalue_plots"]][["mfp_plot_over"]]
ont_up[["pvalue_plots"]][["bpp_plot_over"]]
ont_up[["pvalue_plots"]][["kegg_plot_over"]]
ont_up[["pvalue_plots"]][["reactome_plot_over"]]
ont_up[["pvalue_plots"]][["hp_plot_over"]]

ont_down <- simple_gprofiler(down)
ont_down[["pvalue_plots"]][["mfp_plot_over"]]
ont_down[["pvalue_plots"]][["bpp_plot_over"]]
ont_down[["pvalue_plots"]][["kegg_plot_over"]]
ont_down[["pvalue_plots"]][["reactome_plot_over"]]
ont_down[["pvalue_plots"]][["hp_plot_over"]]
```

```{r cp6}
ont_up <- simple_clusterprofiler(sig_genes=up, de_table=de,
                                 do_david=TRUE, david_user="abelew@umd.edu",
                                 fc_column="deseq_logfc", orgdb="org.Hs.eg.db")
ont_up$plots$ego_sig_mf
ont_up$plots$ego_sig_bp
ont_up$plots$dot_sig_mf
ont_up$plots$dot_sig_bp
ont_up$plots$map_sig_mf
ont_up$plots$map_sig_bp

ont_down <- simple_clusterprofiler(sig_genes=down, de_table=de,
                                   do_david=TRUE, david_user="abelew@umd.edu",
                                   fc_column="deseq_logfc", orgdb="org.Hs.eg.db")
## No significant hits going down
ont_down$plots$ego_all_mf
ont_down$plots$ego_all_bp
ont_down$plots$dot_all_mf
ont_down$plots$dot_all_bp
ont_down$plots$map_all_mf
ont_down$plots$map_all_bp
```

### M LP vs M NS

```{r}
table <- "M_LP_vs_M_NS"
up <- pat_gs_sva_sig[["deseq"]][["ups"]][[table]]
down <- pat_gs_sva_sig[["deseq"]][["downs"]][[table]]
de <- pat_gs_sva_tables[["data"]][[table]]
```

```{r gp7, eval=FALSE}
ont_up <- simple_gprofiler(up)
ont_up[["pvalue_plots"]][["mfp_plot_over"]]
ont_up[["pvalue_plots"]][["bpp_plot_over"]]
ont_up[["pvalue_plots"]][["kegg_plot_over"]]
ont_up[["pvalue_plots"]][["reactome_plot_over"]]
ont_up[["pvalue_plots"]][["hp_plot_over"]]

ont_down <- simple_gprofiler(down)
ont_down[["pvalue_plots"]][["mfp_plot_over"]]
ont_down[["pvalue_plots"]][["bpp_plot_over"]]
ont_down[["pvalue_plots"]][["kegg_plot_over"]]
ont_down[["pvalue_plots"]][["reactome_plot_over"]]
ont_down[["pvalue_plots"]][["hp_plot_over"]]
```

```{r cp7}
ont_up <- simple_clusterprofiler(sig_genes=up, de_table=de,
                                 do_david=TRUE, david_user="abelew@umd.edu",
                                 fc_column="deseq_logfc", orgdb="org.Hs.eg.db")
ont_up$plots$ego_sig_mf
ont_up$plots$ego_sig_bp
ont_up$plots$dot_sig_mf
ont_up$plots$dot_sig_bp
ont_up$plots$map_sig_mf
ont_up$plots$map_sig_bp

ont_down <- simple_clusterprofiler(sig_genes=down, de_table=de,
                                   do_david=TRUE, david_user="abelew@umd.edu",
                                   fc_column="deseq_logfc", orgdb="org.Hs.eg.db")
## No significant hits going down
ont_down$plots$ego_all_mf
ont_down$plots$ego_all_bp
ont_down$plots$dot_all_mf
ont_down$plots$dot_all_bp
ont_down$plots$map_all_mf
ont_down$plots$map_all_bp
```

### M LA vs M NS

```{r}
table <- "M_LA_vs_M_NS"
up <- pat_gs_sva_sig[["deseq"]][["ups"]][[table]]
down <- pat_gs_sva_sig[["deseq"]][["downs"]][[table]]
de <- pat_gs_sva_tables[["data"]][[table]]
```

```{r gp8, eval=FALSE}
ont_up <- simple_gprofiler(up)
ont_up[["pvalue_plots"]][["mfp_plot_over"]]
ont_up[["pvalue_plots"]][["bpp_plot_over"]]
ont_up[["pvalue_plots"]][["kegg_plot_over"]]
ont_up[["pvalue_plots"]][["reactome_plot_over"]]
ont_up[["pvalue_plots"]][["hp_plot_over"]]

ont_down <- simple_gprofiler(down)
ont_down[["pvalue_plots"]][["mfp_plot_over"]]
ont_down[["pvalue_plots"]][["bpp_plot_over"]]
ont_down[["pvalue_plots"]][["kegg_plot_over"]]
ont_down[["pvalue_plots"]][["reactome_plot_over"]]
ont_down[["pvalue_plots"]][["hp_plot_over"]]
```

```{r cp8}
ont_up <- simple_clusterprofiler(sig_genes=up, de_table=de,
                                 do_david=TRUE, david_user="abelew@umd.edu",
                                 fc_column="deseq_logfc", orgdb="org.Hs.eg.db")
ont_up$plots$ego_sig_mf
ont_up$plots$ego_sig_bp
ont_up$plots$dot_sig_mf
ont_up$plots$dot_sig_bp
ont_up$plots$map_sig_mf
ont_up$plots$map_sig_bp

ont_down <- simple_clusterprofiler(sig_genes=down, de_table=de,
                                   do_david=TRUE, david_user="abelew@umd.edu",
                                   fc_column="deseq_logfc", orgdb="org.Hs.eg.db")
ont_down$plots$ego_all_mf
ont_down$plots$ego_all_bp
ont_down$plots$dot_all_mf
ont_down$plots$dot_all_bp
ont_down$plots$map_all_mf
ont_down$plots$map_all_bp
```

## M vs GM

```{r}
table <- "GM_NS_vs_M_NS"
up <- pat_gs_sva_sig[["deseq"]][["ups"]][[table]]
down <- pat_gs_sva_sig[["deseq"]][["downs"]][[table]]
de <- pat_gs_sva_tables[["data"]][[table]]
```

```{r cp9}
ont_up <- simple_clusterprofiler(sig_genes=up, de_table=de,
                                 do_david=TRUE, david_user="abelew@umd.edu",
                                 fc_column="deseq_logfc", orgdb="org.Hs.eg.db")
ont_up$plots$ego_sig_mf
ont_up$plots$ego_sig_bp
ont_up$plots$dot_sig_mf
ont_up$plots$dot_sig_bp
ont_up$plots$map_sig_mf
ont_up$plots$map_sig_bp

ont_down <- simple_clusterprofiler(sig_genes=down, de_table=de,
                                   do_david=TRUE, david_user="abelew@umd.edu",
                                   fc_column="deseq_logfc", orgdb="org.Hs.eg.db")
ont_down$plots$ego_all_mf
ont_down$plots$ego_all_bp
ont_down$plots$dot_all_mf
ont_down$plots$dot_all_bp
ont_down$plots$map_all_mf
ont_down$plots$map_all_bp
```




```{r saveme, eval=FALSE}
pander::pander(sessionInfo())
message(paste0("This is hpgltools commit: ", get_git_commit()))
this_save <- paste0(gsub(pattern="\\.Rmd", replace="", x=rmd_file), "-v", ver, ".rda.xz")
message(paste0("Saving to ", this_save))
tmp <- sm(saveme(filename=this_save))
```


```{r loadme, eval=FALSE}
loadme(filename=this_save)
```
